---
title: "Understand BQ exports"
author: "Eric Schares"
format: html
editor: visual
---

# Process BQ Exports - Gold and Hybrid Journals, using Full APC Dataset, v2

```{r}
library(tidyverse)
library(readxl)
library(scales)
```

```{r}
here::here()
getwd()
```

## Load GBQ export data

```{r}
data <- read_csv('BQ_export_Walden_fullAPCdataset_v5_297167rows.csv')

head(data)
```

```{r}
str(data)
```

Full dataset export returns 12,348 distinct unique_IDs

```{r}
n_distinct(data$unique_ID)
```

```{r}
ggplot(filter(data, oa_type=='ALL'), aes(x=year, y=DOIs)) + 
  geom_col() +
  #scale_y_continuous(labels = comma, breaks = c(0, 25000, 50000, 75000, 100000, 125000, 150000))
  scale_y_continuous(labels = comma) +
  labs(title = 'Number of DOIs by Year')
```

```{r}
data |> 
  filter(oa_type == 'ALL') |> 
  group_by(year) |> 
  summarize(DOIs = sum(DOIs))
```

## Convert GBQ output from long to wide format. One journal-year per row with OA status breakdowns in the columns

```{r}
data_wide <- data |> 
  pivot_wider(id_cols = c('ISSN_L','year'),
                         names_from = oa_type,
                         values_from = DOIs,
                         values_fill = 0,   # fill missing cells with 0
                         values_fn = first) # if there are duplicate ISSN_L-years, keep the first one

view(data_wide)
```

```{r}
# How many observations of an ISSN do we have? Should be 7 (2019-2025)
data_wide |> 
  count(ISSN_L, sort = TRUE)
```

### Construct and plot new columns of `APC-able` documents

```{r}
data_wide <- data_wide |> 
  mutate(goldANDdiamond = gold + diamond) |> 
  mutate(goldDiamondHybrid = gold + diamond + hybrid)
```

ISSN-years that lie along the red 45' line have all their results as Gold, Diamond, or Hybrid

Those at the bottom have Green, Bronze, or Closed

```{r}
ggplot(data_wide, aes(x=ALL, y=goldDiamondHybrid)) + 
  geom_point() + 
  geom_abline(intercept = 0, slope = 1, col = 'red', linetype = 'dashed') +
  labs(title = 'OA type breakdown for each journal-year combination',
       x = 'All articles',
       y = 'Gold + Diamond+ Hybrid articles')
```

## Make the delta column

```{r}
with_delta <- data_wide |> 
  mutate(delta = ALL - goldDiamondHybrid) |> 
  arrange(desc(delta))
```

## Prepeare to filter down journal pub counts to what's in the APC Dataset

```{r}
with_delta <- with_delta |> 
  mutate(issn_year = paste(ISSN_L, year, sep='_'))

head(with_delta)
```

## Load the APC Dataset, v2

```{r}
# csv version
apc_dataset <- read_csv('apc_list_2025_5_analysis_v5_with_flatfees.csv') |> 
  select(journal_normalized, issn_l, publisher, oa_status, apc_year, apc_usd) |> 
  mutate(issn_year = paste(issn_l, apc_year, sep='_'))
head(apc_dataset)
str(apc_dataset)

```

### Filter down to dataset coverage

```{r}
with_delta <- with_delta %>%
  mutate(flag = if_else(
    issn_year %in% apc_dataset$issn_year, 
    1, 
    0
  ))

with_delta <- with_delta |> 
  filter(flag == 1)

head(with_delta)
```

## Combine with APC dataset, bring in APC values

```{r}
combined <- left_join(with_delta, apc_dataset, by = join_by('issn_year'))
combined
```

```{r}
# change NA to 0 in APC prices
combined$apc_usd <- as.numeric(as.character(combined$apc_usd))   # first convert apc_usd from character to numeric
combined$apc_usd[is.na(combined$apc_usd)] <- 0

combined
```

Calculate APC-able volume \* APC

```{r}
combined <- combined |> 
  mutate(volumeXapc = goldDiamondHybrid * apc_usd)
combined
```

```{r}
#write.csv(combined, '12348_Gold&Hybrid_Walden_with_delta_filteredtoAPCdataset_withflatfees_and_APCs.csv', row.names=FALSE)

#combined <- read.csv('12348_Gold&Hybrid_Walden_with_delta_filteredtoAPCdataset_withflatfees_and_APCs.csv')
```

```{r}
combined |> 
  filter(oa_status == 'Gold OA', 
         ALL > 0
         ) |> 
  summarize('Gold OA with at least 1 article' = n(),
            'Article count' = sum(goldDiamondHybrid))
```

```{r}
combined |> 
  filter(oa_status == 'Hybrid OA', 
         ALL > 0
         ) |> 
  summarize('Hybrid OA with at least 1 article' = n(),
            'Article count' = sum(goldDiamondHybrid))
```

```{r}
combined |> 
  filter(oa_status == 'Hybrid OA', 
         ALL > 0,
         goldDiamondHybrid > 0
         ) |> 
  summarize('Hybrid OA with at least 1 APC-able article' = n(),
            'Article count' = sum(goldDiamondHybrid))
```

```{r}
  ggplot(combined, aes(x=volumeXapc, y=apc_usd, color=as_factor(year))) + 
    geom_point() +
    labs(title = 'Volume vs. APC price, 12348 Gold&Hybrid journals, all years',
         x = 'Volume X APC ($)',
         y = 'APC ($)') +
  scale_x_continuous(labels = comma)
  #facet_wrap(~oa_status)

```

```{r}
  ggplot(filter(combined, apc_year==2025), aes(x=goldDiamondHybrid, y=apc_usd, color=as_factor(publisher))) + 
    geom_point() +
    labs(title = 'Volume vs. APC price, Gold&Hybrid journals, 2025 only',
         x = 'Volume (num pubs), log10',
         y = 'APC ($)') +
  scale_x_continuous(labels = comma, transform = 'log10') +
  facet_wrap(~oa_status)

```

```{r}
fig <- ggplot(combined, aes(fill = publisher, x = apc_year, y = volumeXapc)) + 
  geom_bar(position = 'stack', stat = 'identity') + 
  scale_y_continuous(labels = scales::unit_format(unit = "B", scale = 1e-9)) +
  #scale_y_continuous(labels = label_number(scale = 1e-9, suffix = 'B')) +      # could also do
  
  scale_x_continuous(breaks=seq(2019, 2025, 1)) + 
  labs(title = '12348 journals, OpenAlex Walden',
       y = 'USD ($)',
       x = 'Year')

fig

ggsave(plot = fig, filename = "Stacked_bar_chart_by_publisher_year.png")
```

------------------------------------------------------------------------

### Look at OA_status growth of \$ and volume, all Publishers combined

```{r}
overall <- combined |> 
  group_by(oa_status, apc_year) |> 
  summarize(sum = sum(volumeXapc),
            APCable = sum(goldDiamondHybrid),
            .groups = 'drop') |> 
  mutate(avgAPC = sum / APCable)

overall
```

#### OA_status publication volume growth, 2019-2025

```{r}
overall |> 
  filter(apc_year == 2019 | apc_year == 2025) |> 
  pivot_wider(id_cols = c(oa_status),
              names_from = 'apc_year',
              names_prefix = 'year_',
              values_from = 'APCable') |> 
  mutate(
    growth_multiplier = round(year_2025 / year_2019, 5),
    growth_percentage = ((year_2025 - year_2019) / year_2019) * 100
  )
```

#### Publication volume growth by publisher, all OA statuses together

```{r}
combined |> 
  filter(apc_year == 2019 | apc_year == 2025) |> 
  group_by(publisher, apc_year) |> 
  
  summarize(sum = sum(volumeXapc),
            APCable = sum(goldDiamondHybrid),
            .groups = 'drop') |> 
  
  pivot_wider(id_cols = c(publisher),
              names_from = 'apc_year',
              names_prefix = 'year_',
              values_from = 'APCable') |> 
  
  mutate(
    growth_multiplier = round(year_2025 / year_2019, 5),
    growth_percentage = ((year_2025 - year_2019) / year_2019) * 100
  )
```

```{r}
combined |> 
  filter(publisher == 'WILEY') |> 
  group_by(publisher, oa_status, apc_year) |> 
  
  
  summarize(sum = sum(volumeXapc),
            APCable = sum(goldDiamondHybrid),
            .groups = 'drop') |> 
  
  ggplot(aes(x = apc_year, y = sum)) + 
    geom_point(aes(color = oa_status)) + 
  labs(title = 'Wiley spend amount over years')
  

  
```

### Spend (\$) growth by Publisher, year, and OA status

```{r}
grouped_summary <- combined |> 
  group_by(publisher, oa_status, apc_year) |> 
  summarize(sum = sum(volumeXapc),
            APCable = sum(goldDiamondHybrid),
            .groups = 'drop') |> 
  mutate(avgAPC = sum / APCable)

grouped_summary

write.csv(grouped_summary, '12348_journals_bypublisher_withflatfees_money_summary.csv', row.names=FALSE)
```

### Spend (\$) growth by year and OA status, all publishers together

```{r}
# by  year and oa_status breakdown, Table 2
grouped_summary |> 
  group_by(apc_year, oa_status) |> 
  summarize(estimate = round(sum(sum)/1e6,2)) |>    # in Millions
  #mutate(estimate = num(estimate, sigfig = 5, notation = "si"))
  as.tibble()
```

### Spend (\$) sum by OA status, all publishers and years together

```{r}
# by oa_status breakdown, sum of all years
grouped_summary |> 
  group_by(oa_status) |> 
  summarize(estimate = round(sum(sum)/1e6,2)) |>    # in Millions
  #mutate(estimate = num(estimate, sigfig = 5, notation = "si"))
  as.tibble()
```

## By Year Summary, all publishers and oa_status together

```{r}
combined |> 
  group_by(apc_year, oa_status) |> 
  summarize(sum_in_M_USD = sum(volumeXapc) / 1e6,
            .groups = 'drop')
```

## Growth percentages of article output from 2019-2025 by publisher and oa_status

```{r}
df_growth <- grouped_summary |> 
  filter(apc_year == 2019 | apc_year == 2025) |> 
  pivot_wider(id_cols = c(publisher, oa_status),
              names_from = 'apc_year',
              names_prefix = 'year_',
              values_from = 'APCable') |> 
  mutate(
    growth_multiplier = round(year_2025 / year_2019, 5),
    growth_percentage = ((year_2025 - year_2019) / year_2019) * 100
  )

df_growth

```

## Article output summary by publisher, year, and oa_status

```{r}
df_article_summary <- combined |> 
  group_by(publisher, apc_year, oa_status) |> 
  
  summarize(APCable = sum(goldDiamondHybrid),
            .groups = 'drop') |> 
  
  pivot_wider(id_cols = c(publisher, apc_year),
              names_from = 'oa_status',
              #names_prefix = 'year_',
              values_from = 'APCable', 
              values_fill = 0)
  

  # mutate(
  #   growth_multiplier = round(year_2025 / year_2019, 5),
  #   growth_percentage = ((year_2025 - year_2019) / year_2019) * 100
  # )

df_article_summary

#write.csv(df_article_summary, 'article_counts_by_publisher_year_OAtype.csv', row.names=FALSE)

```

```{r}
# Calculate YoY percentage increase
df_article_summary |>
  
  # sort years first so lag() pulls the true prior year.
  arrange(publisher, apc_year) |> 
  group_by(publisher) |> 
  
  # lag() looks at row directly above the current one
  mutate(
    `Gold OA YoY %`   = (`Gold OA` - lag(`Gold OA`)) / lag(`Gold OA`) * 100,
    `Gold OA YoY multiplier`   = (`Gold OA` / lag(`Gold OA`)),
    
    `Hybrid OA YoY %` = (`Hybrid OA` - lag(`Hybrid OA`)) / lag(`Hybrid OA`) * 100,
    `Hybrid OA YoY multiplier` = (`Hybrid OA` / lag(`Hybrid OA`))
  ) |> 
  
  ungroup() |> 
  
  write.csv('article_counts_by_publisher_year_OAtype_withpercents.csv', row.names=FALSE)
```
