---
title: "roara_apc_paper_without_inflation_adjustment"
author: "Lisa Matthias"
date: "2026-07-29"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

```{r include=FALSE}
library(tidyverse)
```

```{r main-data, include=FALSE}
apc_raw <- read_csv("./raw_data/apc_list_2025_5_analysis_copy.csv")

apc_aggregate <- read_csv("./data/12348_journals_bypublisher_withflatfees_money_summary.csv")

journals_pub_counts_apc <- read_csv(
  "./data/12348_Gold&Hybrid_Walden_with_delta_filteredtoAPCdataset_withflatfees_and_APCs.csv"
)
```

```{r colors, include=FALSE}
# Palette OA Types
pal_oa_types <- c("Gold OA" = "#D6A94E", "Hybrid OA" = "#B85428")

# Palette Publishers
pal_top5 <- c(
  "Elsevier"        = "#ff7f0f",  
  "Springer Nature" = "#00316e",   
  "MDPI"            = "#697754",   
  "Wiley"           = "#b680bf",   
  "Frontiers"       = "#069494"    
)

pal_other <- c(
  "OUP"        = "#205f4b", 
  "ACS"        = "#7B7BE8",
  "PLOS"       = "#7e1a2f",
  "CUP"        = "#ee8577",
  "IOP"        = "#4A8FBF",
  "Sage"       = "#B3B177",
  "IEEE"       = "#2E4057",
  "De Gruyter" = "#a06177",
  "EDP"        = "#D58E40"
)
```

The main text of our manuscript reports APC spend and revenue figures adjusted for inflation to 2025 USD using CPI Advanced Economies (see Methods). This supplement provides the corresponding unadjusted, nominal figures—the actual list-price values for each year, without adjustment. Table and figure numbers below correspond to their inflation-adjusted counterparts in the main text.

### Figure S2.2 Estimate of annual APC spend (in USD) per OA type

```{r figS2-2-apc-spend-gold-hybrid-chicklet-no-adjustment,echo=FALSE, fig.dpi=300}
library(ggchicklet)   
library(scales)
library(ggtext)

df_spend_by_publisher_oa_type <- apc_aggregate %>%
  filter(oa_status %in% c("Gold OA", "Hybrid OA")) 

# Total revenue per year, per OA type (summed across all publishers)
df_year_oa_type <- df_spend_by_publisher_oa_type %>%
  group_by(apc_year, oa_status) %>%
  summarise(total = sum(sum, na.rm = TRUE), .groups = "drop")

# Lock stacking order explicitly: Gold OA bottom, Hybrid OA top
df_year_oa_type <- df_year_oa_type %>%
  mutate(oa_status = factor(oa_status, levels = c("Gold OA", "Hybrid OA")))

# Compute cumulative stack positions manually so labels can be placed precisely
df_year_oa_type <- df_year_oa_type %>%
  arrange(apc_year, oa_status) %>%
  group_by(apc_year) %>%
  mutate(
    ymax        = cumsum(total),
    ymin        = ymax - total,
    grand_total_bar = sum(total),
    seg_share   = total / grand_total_bar
  ) %>%
  ungroup()

df_year_oa_type <- df_year_oa_type %>%
  mutate(is_thin = seg_share < 0.12)

# Custom label: one decimal for billions, whole numbers for millions
format_mixed <- function(x) {
  if_else(
    x >= 1e9,
    paste0("$", round(x / 1e9, 1), "B"),
    paste0("$", round(x / 1e6, 0), "M")
  )
}

# Endpoint labels: total per segment, placed inside/at top of each chicklet
df_year_oa_type <- df_year_oa_type %>%
  mutate(label = format_mixed(total))

# Grand total per year (for the overall annotation above each bar)
df_total <- df_year_oa_type %>%
  group_by(apc_year) %>%
  summarise(grand_total = sum(total, na.rm = TRUE)) %>%
  mutate(label = format_mixed(grand_total))

# Plot
ggplot(df_year_oa_type, aes(x = factor(apc_year), y = total, fill = oa_status)) +
  geom_chicklet(
    radius = grid::unit(3, "pt"),
    width = 0.7
  ) +
  geom_text(
    aes(label = label, y = (ymin + ymax) / 2, vjust = if_else(is_thin, 0.7, 0.5)),
    color = "white", fontface = "bold", size = 3.4
  ) +
  geom_text(
    data = df_total,
    aes(x = factor(apc_year), y = grand_total, label = label),
    inherit.aes = FALSE,
    vjust = -0.6, fontface = "bold", size = 4, color = "black"
  ) +
  scale_fill_manual(values = pal_oa_types, name = NULL) +
  scale_y_continuous(
    labels = label_number(scale_cut = cut_short_scale(), prefix = "$"),
    expand = expansion(mult = c(0, 0.12))
  ) +
  labs(
    title = "Estimate of annual APC spend (in USD) by OA type",
    caption = "Some publishers are not covered in every year (De Gruyter and OUP begin in 2021, IOP is missing 2020, and Sage's 2019 and 2025 data are gold-only and hybrid-only respectively).",
    x = NULL, y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor    = element_blank(),
    panel.grid.major.y = element_blank(),
    axis.text.x  = element_text(size = 12, face = "bold", color = "black"),
    axis.text.y  = element_blank(),
    axis.line.x.bottom  = element_line(color = "black", linewidth = 0.4),
    legend.position = "top",
    legend.justification = "left",
    plot.title    = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(color = "black", size = 11),
    plot.caption = element_textbox_simple(
      size = 9,
      color = "black",
      lineheight = 1.3,
      margin = margin(t = 12)
    )
  )

ggsave("figures/figS2_2_apc_spend_chicklet_not_adjusted.png", width = 7, height = 5, dpi = 300, bg = "white")
```

## Publishers

```{r function-clean-publisher-labels,echo=FALSE}
# Clean publisher labels
clean_label <- function(x) {
  x %>%
    str_replace_all("_", " ") %>%
    str_to_title() %>%
    str_replace("Acs", "ACS") %>%
    str_replace("Cambridge University Press - Cup", "CUP") %>%
    str_replace("Edp", "EDP") %>%
    str_replace("Ieee", "IEEE") %>%
    str_replace("Iop", "IOP") %>%
    str_replace("Mdpi", "MDPI") %>%
    str_replace("Oup", "OUP") %>%
    str_replace("Plos", "PLOS")
}
```

```{r figS2-3-plot-function,echo=FALSE}
make_fig3 <- function(data, ends_data, title, y_breaks, pal, use_repel = TRUE) {
  lab_layer <- if (use_repel) {
    geom_text_repel(
      data = ends_data,
      aes(x = x_lab, y = revenue_m, label = lab),
      hjust              = 0,
      direction          = "y",
      size               = 3.1,
      fontface           = "bold",
      segment.colour     = "grey75",
      segment.size       = 0.25,
      min.segment.length = 0.4,
      box.padding        = 0.25,
      show.legend        = FALSE
    )
  } else {
    geom_text(
      data = ends_data,
      aes(x = x_lab, y = revenue_m, label = lab),
      hjust       = 0,
      size        = 3.1,
      fontface    = "bold",
      show.legend = FALSE
    )
  }

  ggplot(data, aes(apc_year, revenue_m, colour = publisher)) +

    # hand-drawn gridlines: stop at 2025, don't run into the label gutter
    geom_segment(
      data = tibble(y = y_breaks),
      aes(x = 2019, xend = 2025, y = y, yend = y),
      inherit.aes = FALSE,
      colour = "grey92", linewidth = 0.4
    ) +

    geom_line(linewidth = 0.9) +
    geom_point(size = 1.6) +
    lab_layer +

    scale_x_continuous(
      breaks = 2019:2025,
      expand = expansion(mult = c(0.03, 0.42))   # gutter for the labels
    ) +
    scale_y_continuous(
      breaks = y_breaks,
      labels = label_dollar(suffix = "M", big.mark = ",", accuracy = 1),
      expand = expansion(mult = c(0.05, 0.08))
    ) +
    scale_colour_manual(values = pal, drop = TRUE) +
    labs(title = title, x = NULL, y = NULL) +
    theme_minimal(base_size = 12) +
    theme(
      legend.position = "none",
      panel.grid      = element_blank(),
      plot.title      = element_text(face = "bold", size = 12,
                                     margin = margin(b = 6))
    )
}
```

### Figure S2.3 Estimate of annual APC spend (in USD) by publisher.

```{r figS2-3-publishers-apc-spend,echo=FALSE,fig.width = 6.5, fig.height = 7.5, fig.dpi=300}
library(scales)
library(patchwork)
library(ggrepel)

# Data 
rev_by_pub <- apc_aggregate %>%
  filter(oa_status %in% c("Gold OA", "Hybrid OA")) %>%
  group_by(publisher, apc_year) %>%
  summarise(revenue_m = sum(sum, na.rm = TRUE) / 1e6, .groups = "drop") %>%
  mutate(publisher = clean_label(publisher)) %>%
  filter(!(publisher == "IOP" & apc_year == 2020)) # no data for 2020

# Rank publishers by total revenue across all years
top5 <- rev_by_pub %>%
  group_by(publisher) %>%
  summarise(tot = sum(revenue_m), .groups = "drop") %>%
  slice_max(tot, n = 5) %>%
  pull(publisher)

rev_by_pub <- rev_by_pub %>%
  mutate(group = if_else(publisher %in% top5, "Top 5 publishers", "Other publishers"))

# End-of-line labels, pinned just right of the last data point
ends <- rev_by_pub %>%
  group_by(publisher, group) %>%
  filter(apc_year == max(apc_year)) %>%
  ungroup() %>%
  mutate(
    lab = paste0(publisher, "  $",
                 formatC(round(revenue_m), format = "d", big.mark = ","),
                 "M"),
    x_lab = 2025.25
  )

# Plot 
p_top5 <- make_fig3(
  filter(rev_by_pub, group == "Top 5 publishers"),
  filter(ends,       group == "Top 5 publishers"),
  "Top 5 publishers",
  y_breaks  = seq(0, 1000, by = 250),
  pal       = pal_top5,
  use_repel = FALSE
)

p_next9 <- make_fig3(
  filter(rev_by_pub, group == "Other publishers"),
  filter(ends,       group == "Other publishers"),
  "Next 9 publishers",
  y_breaks  = seq(0, 120, by = 30),
  pal = pal_other,
  use_repel = TRUE
)

publishers_apc_spend_total_not_adjusted <- p_top5 / p_next9 +
  plot_annotation(
    title    = "Estimate of annual APC spend (in USD) by publisher",
    caption = "Sage's 2019 and 2025 figures reflect gold-only and hybrid-only data respectively.",
    theme    = theme(
      plot.title    = element_text(face = "bold", size = 15),
      plot.subtitle = element_text(colour = "black", size = 11,
                                   margin = margin(b = 10)),
      plot.caption = element_textbox_simple(
      size = 9,
      color = "black",
      lineheight = 1.3,
      margin = margin(t = 12)
    )
    )
  )

ggsave("figS2_3_publishers_apc_spend_total_not_adjusted.png", publishers_apc_spend_total_not_adjusted,
       width = 6.5, height = 7.5, dpi = 300, bg = "white")

publishers_apc_spend_total_not_adjusted
```

### Figure S2.4 Estimate of annual APC revenue (in USD) by publisher and OA type.

```{r figS2-4-publishers-apc-spend-oa-type-not-adjusted, echo=FALSE, include=FALSE}
library(ggtext)
library(scales)
library(ggh4x)
library(grid)

# Data
df_spend_fig4_not_adjusted <- df_spend_by_publisher_oa_type %>%
    mutate(sum = if_else(sum == 0, NA_real_, sum)) # so that Sage 2025 doesn't print as $0

# Explicit panel order (by apc spend in 2025)
pub_order <- c(
  "ELSEVIER", "SPRINGER_NATURE", "MDPI",
  "WILEY", "FRONTIERS", "OUP",
  "ACS", "PLOS", "CAMBRIDGE_UNIVERSITY_PRESS_-_CUP",
  "IOP", "SAGE", "IEEE",
  "DE_GRUYTER", "EDP"
)

# Fail on any name mismatch rather than silently dropping a panel
stopifnot(length(setdiff(pub_order, unique(df_spend_fig4_not_adjusted$publisher))) == 0)
stopifnot(length(setdiff(unique(df_spend_fig4_not_adjusted$publisher), pub_order)) == 0)

pub_labels <- setNames(clean_label(pub_order), pub_order)

df_spend_fig4_ordered_not_adjusted <- df_spend_fig4_not_adjusted %>%
  mutate(publisher = factor(publisher, levels = pub_order))

# Peaks in panel order to check these clear the row ceilings set below
df_spend_fig4_ordered_not_adjusted %>%
  group_by(publisher) %>%
  summarise(peak = max(sum, na.rm = TRUE), .groups = "drop") %>%
  arrange(publisher) %>%
  print(n = Inf)

# Endpoint labels
df_ends <- df_spend_fig4_ordered_not_adjusted %>%
  filter(apc_year == 2025) %>%
  mutate(label = label_number(scale_cut = cut_short_scale(),
                              prefix = "$", accuracy = 1)(sum))

## Sage has no 2025 gold data; use its 2024 gold endpoint instead
df_ends_sage <- df_spend_fig4_ordered_not_adjusted %>%
  filter(apc_year == 2024, publisher == "SAGE", oa_status == "Gold OA") %>%
  mutate(label = label_number(scale_cut = cut_short_scale(),
                              prefix = "$", accuracy = 1)(sum))

df_ends <- bind_rows(df_ends, df_ends_sage) %>%
  mutate(publisher = factor(publisher, levels = pub_order))

# Per-row y-axis ceilings 
nc <- 3   

# Row 1: Elsevier, Springer Nature, MDPI
# Row 2: Wiley, FRONTIERS, OUP
# Row 3: ACS, PLOS, CUP
# Row 4: IOP, SAGE, IEEE
# Row 5: De Gruyter, EDP
row_specs <- list(
  list(limit = 700e6, breaks = c(0, 200e6, 400e6, 600e6)),
  list(limit = 320e6, breaks = c(0, 100e6, 200e6, 300e6)),
  list(limit =  60e6, breaks = c(0,  20e6,  40e6,  60e6)),
  list(limit =  60e6, breaks = c(0,  20e6,  40e6,  60e6)),
  list(limit =  8e6, breaks = c(0,   4e6,   8e6,  12e6))
)

y_scales <- vector("list", length(pub_order))

for (row_num in seq_along(row_specs)) {
  positions <- ((row_num - 1) * nc + 1):(row_num * nc)
  positions <- positions[positions <= length(pub_order)]
  spec <- row_specs[[row_num]]
  for (pos in positions) {
    y_scales[[pos]] <- scale_y_continuous(
      breaks = spec$breaks,
      labels = label_number(scale_cut = cut_short_scale(), prefix = "$"),
      limits = c(0, spec$limit),
      expand = expansion(mult = c(0.05, 0.15))
    )
  }
}

# OUP sits in row 2 (with Wiley and FRONTIERS) but belongs on a 60M axis
oup_pos <- which(pub_order == "OUP")
y_scales[[oup_pos]] <- scale_y_continuous(
  breaks = c(0, 20e6, 40e6, 60e6),
  labels = label_number(scale_cut = cut_short_scale(), prefix = "$"),
  limits = c(0, 60e6),
  expand = expansion(mult = c(0.05, 0.15))
)

# Plot
p <- ggplot(
  df_spend_fig4_ordered_not_adjusted,
  aes(
    x = apc_year,
    y = sum,
    colour = oa_status,
    group = oa_status
  )
) +
  geom_line(linewidth = 0.85) +
  geom_point(size = 3, shape = 16) +
  geom_text(
    data = df_ends,
    aes(label = label),
    hjust = -0.15,
    size = 5,
    fontface = "bold",
    show.legend = FALSE
  ) +
  facet_wrap2(
    ~ publisher,
    scales = "free_y",
    ncol = nc,
    labeller = as_labeller(pub_labels)
  ) +
  facetted_pos_scales(y = y_scales) +
  scale_colour_manual(values = pal_oa_types, name = NULL) +
  scale_x_continuous(
    breaks = c(2019, 2022, 2025),
    expand = expansion(mult = c(0.05, 0.42))   # room for the endpoint labels
  ) +
  labs(
    title    = "Estimate of annual APC spend (in USD) by publisher and OA type",
    caption  = "Panels are ordered by total 2025 APC spend Note that the y-axis scale differs between panels.",
    x = NULL,
    y = NULL
  ) +
  theme_minimal() +
  theme(
    panel.grid.major.x   = element_blank(),
    panel.grid.major.y   = element_line(colour = "grey92", linewidth = 0.3),
    panel.grid.minor     = element_blank(),
    strip.text           = element_text(
      face = "bold",
      size = 14,
      colour = "black"
    ),
    legend.position      = "top",
    legend.justification = "left",
    legend.text          = element_text(size = 16),
    legend.key.width     = unit(2, "lines"),
    axis.text.x          = element_text(size = 14),
    axis.text.y          = element_text(size = 14),
    plot.title           = element_text(face = "bold", size = 18),
    plot.subtitle        = element_text(colour = "black", size = 16),
    plot.caption         = element_text(
      size = 14,
      colour = "black",
      margin = margin(t = 15),
      hjust = 0
    ),
    plot.caption.position = "plot",
    plot.margin = margin(b = 15)
  )
```

```{r figS2-4-publishers-apc-spend-oa-type-not-adjusted-plot, echo=FALSE, fig.width=10, fig.height=13}
print(p)

png(
  "figures/figS2_4_apc_spend_publisher_oa_type_not_adjusted.png",
  width = 10 * 300,
  height = 13 * 300,
  res = 300
)
print(p)
grid.lines(
  x  = unit(0.675, "npc"),
  y  = unit(c(0.57, 0.74), "npc"),
  gp = gpar(
    col = "black",
    lwd = 2.5,
    lty = "dashed"
  )
)
dev.off()

```

### Table S2.2 Growth of article output and APC spend from 2019 to 2025 per publisher and OA type.

```{r tableS2-2-spend-growth-not-adjusted, echo=FALSE}
library(kableExtra)

# Spend growth
spend_growth_not_adjusted <- df_spend_by_publisher_oa_type %>%
  filter(apc_year %in% c(2019, 2025))

## growth rate per publisher, split by OA status
growth_by_type_not_adjusted <- spend_growth_not_adjusted %>%
  group_by(publisher, oa_status, apc_year) %>%
  summarise(total = sum(sum), .groups = "drop") %>%
  pivot_wider(names_from = apc_year, values_from = total, names_prefix = "y") %>%
  mutate(growth_pct = round((y2025 - y2019) / y2019 * 100, 1))

## growth rate per publisher, Gold+Hybrid combined
growth_combined_not_adjusted <- spend_growth_not_adjusted %>%
  group_by(publisher, apc_year) %>%
  summarise(total = sum(sum), .groups = "drop") %>%
  pivot_wider(names_from = apc_year, values_from = total, names_prefix = "y") %>%
  mutate(oa_status = "Gold+Hybrid",
         growth_pct = round((y2025 - y2019) / y2019 * 100, 1))

## the "all publishers" row, split by OA status
growth_all_not_adjusted <- spend_growth_not_adjusted %>%
  group_by(oa_status, apc_year) %>%
  summarise(total = sum(sum), .groups = "drop") %>%
  pivot_wider(names_from = apc_year, values_from = total, names_prefix = "y") %>%
  mutate(publisher = "all publishers",
         growth_pct = round((y2025 - y2019) / y2019 * 100, 1))

## the "all publishers" row, Gold+Hybrid combined
growth_all_combined_not_adjusted <- spend_growth_not_adjusted %>%
  group_by(apc_year) %>%
  summarise(total = sum(sum), .groups = "drop") %>%
  pivot_wider(names_from = apc_year, values_from = total, names_prefix = "y") %>%
  mutate(publisher = "all publishers", oa_status = "Gold+Hybrid",
         growth_pct = round((y2025 - y2019) / y2019 * 100, 1))

# spend growth table
table2_spend_growth_not_adjusted <- bind_rows(growth_by_type_not_adjusted, growth_combined_not_adjusted,
                                  growth_all_not_adjusted, growth_all_combined_not_adjusted) %>%
  select(publisher, oa_status, growth_pct) %>%
  pivot_wider(names_from = oa_status, values_from = growth_pct) %>%
  select(publisher, spend_gh = `Gold+Hybrid`, spend_g = `Gold OA`, spend_h = `Hybrid OA`)

# Article output growth
## article-count growth per publisher, split by OA status
article_growth <- journals_pub_counts_apc %>%
  filter(year %in% c(2019, 2025), oa_status %in% c("Gold OA", "Hybrid OA")) %>%
  group_by(publisher, oa_status, year) %>%
  summarise(total = sum(goldDiamondHybrid, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = year, values_from = total, names_prefix = "y") %>%
  mutate(growth_pct = round((y2025 - y2019) / y2019 * 100, 1))

## article-count growth per publisher, Gold+Hybrid combined
article_growth_combined <- journals_pub_counts_apc %>%
  filter(year %in% c(2019, 2025), oa_status %in% c("Gold OA", "Hybrid OA")) %>%
  group_by(publisher, year) %>%
  summarise(total = sum(goldDiamondHybrid, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = year, values_from = total, names_prefix = "y") %>%
  mutate(oa_status = "Gold+Hybrid",
         growth_pct = round((y2025 - y2019) / y2019 * 100, 1))

## the "all publishers" row, split by OA status
article_growth_all <- journals_pub_counts_apc %>%
  filter(year %in% c(2019, 2025), oa_status %in% c("Gold OA", "Hybrid OA")) %>%
  group_by(oa_status, year) %>%
  summarise(total = sum(goldDiamondHybrid, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = year, values_from = total, names_prefix = "y") %>%
  mutate(publisher = "all publishers",
         growth_pct = round((y2025 - y2019) / y2019 * 100, 1))

## the "all publishers" row, Gold+Hybrid combined
article_growth_all_combined <- journals_pub_counts_apc %>%
  filter(year %in% c(2019, 2025), oa_status %in% c("Gold OA", "Hybrid OA")) %>%
  group_by(year) %>%
  summarise(total = sum(goldDiamondHybrid, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = year, values_from = total, names_prefix = "y") %>%
  mutate(publisher = "all publishers", oa_status = "Gold+Hybrid",
         growth_pct = round((y2025 - y2019) / y2019 * 100, 1))

# article grwoth table
table2_article_growth_not_adjusted <- bind_rows(article_growth, article_growth_combined,
                                    article_growth_all, article_growth_all_combined) %>%
  select(publisher, oa_status, growth_pct) %>%
  pivot_wider(names_from = oa_status, values_from = growth_pct) %>%
  select(publisher, articles_gh = `Gold+Hybrid`, articles_g = `Gold OA`, articles_h = `Hybrid OA`)

# Combine article + spend growth into one table
table2_full_not_adjusted <- table2_article_growth_not_adjusted %>%
  left_join(table2_spend_growth_not_adjusted, by = "publisher") %>%
  mutate(publisher = if_else(publisher == "all publishers",
                              "All publishers",
                              clean_label(publisher))) %>%
  arrange(publisher != "All publishers", publisher)

# Add n/a for Sage due to data gaps
table2_full_not_adjusted <- table2_full_not_adjusted %>%
  mutate(across(-publisher, ~ if_else(str_detect(publisher, regex("sage", ignore_case = TRUE)), NA_real_, .x)))

# Save
write_csv(table2_full_not_adjusted, "tableS2_2_growth_full_not_adjusted.csv")

# Print
table2_full_not_adjusted %>%
  mutate(across(-publisher, ~ if_else(is.na(.x), "n/a", sprintf("%+.1f%%", .x)))) %>%
  knitr::kable(
    col.names = c("Publisher", "Gold+Hybrid", "Gold", "Hybrid",
                  "Gold+Hybrid", "Gold", "Hybrid"),
    align = "lrrrrrr",
    caption = "Table S2.2 Growth of article output and APC spend from 2019 to 2025 per publisher and OA type."
  ) %>%
  kable_styling(full_width = FALSE) %>%
  add_header_above(c(
    " " = 1,
    "Number of APC-able publications" = 3,
    "Spend estimate based on APCs in USD (adjusted to 2025 prices)" = 3
  ))
```
