---
title: "roara_apc_paper"
author: "Lisa Matthias"
date: "2026-07-03"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

```{r include=FALSE}
library(tidyverse)
```

```{r main-data, include=FALSE}
apc_raw <- read_csv("./raw_data/apc_list_2025_5_analysis_copy.csv")

apc_aggregate <- read_csv("./data/12348_journals_bypublisher_withflatfees_money_summary.csv")

journals_pub_counts_apc <- read_csv(
  "./data/12348_Gold&Hybrid_Walden_with_delta_filteredtoAPCdataset_withflatfees_and_APCs.csv"
)
```

```{r inflation-adjustment, include=FALSE}
# Rows = target year (what we're converting TO), Cols = base year (converting FROM)
# Cell = multiplier to express a `base` amount in `target` dollars.
# Upper triangle is NA (don't deflate future values back).

infl_matrix <- matrix(
  c(
    1.000,   NA,    NA,    NA,    NA,    NA,   NA,
    1.007, 1.000,   NA,    NA,    NA,    NA,   NA,
    1.038, 1.031, 1.000,   NA,    NA,    NA,   NA,
    1.114, 1.106, 1.073, 1.000,   NA,    NA,   NA,
    1.165, 1.157, 1.122, 1.046, 1.000,   NA,   NA,
    1.196, 1.187, 1.152, 1.073, 1.026, 1.000,  NA,
    1.225, 1.217, 1.180, 1.100, 1.052, 1.025, 1.000
  ),
  nrow = 7, byrow = TRUE,
  dimnames = list(target = 2019:2025, base = 2019:2025)
)

# Look up a single factor, e.g. convert a 2019 amount into 2025 dollars:
infl_matrix["2025", "2019"]   # 1.225

# Factors to convert to 2025 dollars
to_2025 <- c(
  `2019` = 1.225,
  `2020` = 1.217,
  `2021` = 1.180,
  `2022` = 1.100,
  `2023` = 1.052,
  `2024` = 1.025,
  `2025` = 1.000
)
```

```{r colors, include=FALSE}
# Palette OA Types
pal_oa_types <- c("Gold OA" = "#D6A94E", "Hybrid OA" = "#B85428")

# Palette Publishers
pal_top5 <- c(
  "Elsevier"        = "#ff7f0f",  
  "Springer Nature" = "#00316e",   
  "MDPI"            = "#697754",   
  "Wiley"           = "#b680bf",   
  "Frontiers"       = "#069494"    
)

pal_other <- c(
  "OUP"        = "#205f4b", 
  "ACS"        = "#7B7BE8",
  "PLOS"       = "#7e1a2f",
  "CUP"        = "#ee8577",
  "IOP"        = "#4A8FBF",
  "Sage"       = "#B3B177",
  "IEEE"       = "#2E4057",
  "De Gruyter" = "#a06177",
  "EDP"        = "#D58E40"
)
```

# Methods

## Main numbers

Number of records in the dataset

```{r n-records, echo=FALSE}
apc_raw %>% nrow()
```

Number of journals in the dataset

```{r n-journals, echo=FALSE}
length(unique(apc_raw$unique_id_version_2_1))
```

## Currency numbers

Number of flat rate APCs in the dataset

```{r flat-rate-count, echo=FALSE}
flat <- apc_raw %>% filter(type_of_fee == "flat rate")
total <- nrow(flat)
total
```

Original vs. converted currency, flat-rate APCs

```{r currency-original-pct, echo=FALSE}
usd_original <- sum(flat$apc_usd_originalORconverted == "original", na.rm = TRUE)
usd_original_pct <- round(usd_original / total * 100, 1)
usd_original_pct
```

## Fee-type breakdown

```{r fee-type-tally, echo=FALSE}
apc_raw %>% count(type_of_fee)
```

## OA status 

```{r oa-status-tally, echo=FALSE}
apc_raw %>% count(oa_status)
```

Gold OA journal-year combinations with known APCs

```{r gold-known-apc, echo=FALSE}
apc_raw %>%
  filter(
    oa_status == "Gold OA",
    !type_of_fee %in% c("unknown", "apc scale", "liberty apc")
  ) %>%
  nrow()
```

Gold OA journal-year combinations with APC > 0

```{r gold-apc-positive, echo=FALSE}
apc_raw %>%
  filter(
    oa_status == "Gold OA",
    !type_of_fee %in% c("unknown", "apc scale", "liberty apc"),
    suppressWarnings(as.numeric(apc_usd)) > 0
  ) %>%
  nrow()
```

Hybrid OA journal-year combinations with known APCs

```{r hybrid-known-apc, echo=FALSE}
apc_raw %>%
  filter(oa_status == "Hybrid OA", type_of_fee != "unknown") %>%
  nrow()
```

# Results

```{r table1-estimate-annual-APC-spend-cpi, echo=FALSE}
library(kableExtra)

spend_table <- apc_aggregate %>%
  filter(oa_status != "unkown") %>%
  # Sum across publishers -> one row per year per OA status
  group_by(oa_status, apc_year) %>%
  summarise(actual = sum(sum), .groups = "drop") %>%
  # Apply inflation adjustment to 2025 dollars
  mutate(adjusted = actual * to_2025[as.character(apc_year)]) %>%
  # Convert to millions
  mutate(across(c(actual, adjusted), ~ .x / 1e6)) %>%
  # Tidy OA labels for column names
  mutate(oa = case_when(
    str_detect(oa_status, "Gold")   ~ "Gold",
    str_detect(oa_status, "Hybrid") ~ "Hybrid",
    TRUE ~ oa_status
  )) %>%
  select(apc_year, oa, actual, adjusted) %>%
  pivot_wider(
    names_from = oa,
    values_from = c(actual, adjusted),
    names_glue = "{oa}_{.value}"
  ) %>%
  arrange(apc_year)

# Add 2019-2025 total row
totals <- spend_table %>%
  summarise(across(-apc_year, \(x) sum(x, na.rm = TRUE))) %>%
  mutate(apc_year = NA_integer_)

spend_table <- bind_rows(spend_table, totals)

# Order columns to match paper layout: actual Gold/Hybrid, then adjusted Gold/Hybrid
spend_table <- spend_table %>%
  select(apc_year,
         Gold_actual, Hybrid_actual,
         Gold_adjusted, Hybrid_adjusted)

# Display rounded to 1 decimal
spend_table_rounded <- spend_table %>%
  mutate(across(-apc_year, ~ round(.x, 2)))

# Save
spend_table_rounded %>%
  select(apc_year, Gold_actual, Hybrid_actual, Gold_adjusted, Hybrid_adjusted) %>%
  rename(
    `Publication year` = apc_year,
    `Gold (actual list price)`     = Gold_actual,
    `Hybrid (actual list price)`   = Hybrid_actual,
    `Gold (adjusted to 2025)`      = Gold_adjusted,
    `Hybrid (adjusted to 2025)`    = Hybrid_adjusted
  ) %>%
  write_csv("table1_apc_spend.csv")

# Print
spend_table_rounded %>%
  select(apc_year, Gold_actual, Hybrid_actual, Gold_adjusted, Hybrid_adjusted) %>%
  knitr::kable(
    col.names = c("Publication year", "Gold", "Hybrid", "Gold", "Hybrid"),
    align = "lrrrr",
    caption = "Table 1. Estimate of annual APC spend (in millions USD) per OA type for actual APCs paid and adjusted for inflation using CPI Advanced Economies."
  ) %>%
  kable_styling(full_width = FALSE) %>%
  add_header_above(c(
    " " = 1,
    "Spend estimate based on APCs in USD (actual list price)" = 2,
    "Spend estimate based on APCs in USD (adjusted to 2025 prices)" = 2
  ))


```

## OA type

### Figure 2. Estimate of annual APC spend (in USD) per OA type, adjusted for inflation to 2025 USD using CPI Advanced Economies.

```{r fig2-apc-spend-gold-hybrid-chicklet,echo=FALSE, fig.dpi=300}
library(ggchicklet)   
library(scales)
library(ggtext)

df_spend_by_publisher_oa_type_inflation <- apc_aggregate %>%
  filter(oa_status %in% c("Gold OA", "Hybrid OA")) %>%
  mutate(sum_adj = sum * to_2025[as.character(apc_year)])

# Total revenue per year, per OA type (summed across all publishers)
df_year_oa_type <- df_spend_by_publisher_oa_type_inflation %>%
  group_by(apc_year, oa_status) %>%
  summarise(total = sum(sum_adj, na.rm = TRUE), .groups = "drop")

# Lock stacking order explicitly: Gold OA bottom, Hybrid OA top
df_year_oa_type <- df_year_oa_type %>%
  mutate(oa_status = factor(oa_status, levels = c("Gold OA", "Hybrid OA")))

# Compute cumulative stack positions manually so labels can be placed precisely
df_year_oa_type <- df_year_oa_type %>%
  arrange(apc_year, oa_status) %>%
  group_by(apc_year) %>%
  mutate(
    ymax        = cumsum(total),
    ymin        = ymax - total,
    grand_total_bar = sum(total),
    seg_share   = total / grand_total_bar
  ) %>%
  ungroup()

df_year_oa_type <- df_year_oa_type %>%
  mutate(is_thin = seg_share < 0.12)

# Custom label: one decimal for billions, whole numbers for millions
format_mixed <- function(x) {
  if_else(
    x >= 1e9,
    paste0("$", round(x / 1e9, 1), "B"),
    paste0("$", round(x / 1e6, 0), "M")
  )
}

# Endpoint labels: total per segment, placed inside/at top of each chicklet
df_year_oa_type <- df_year_oa_type %>%
  mutate(label = format_mixed(total))

# Grand total per year (for the overall annotation above each bar)
df_total <- df_year_oa_type %>%
  group_by(apc_year) %>%
  summarise(grand_total = sum(total, na.rm = TRUE)) %>%
  mutate(label = format_mixed(grand_total))

# Plot
ggplot(df_year_oa_type, aes(x = factor(apc_year), y = total, fill = oa_status)) +
  geom_chicklet(
    radius = grid::unit(3, "pt"),
    width = 0.7
  ) +
  geom_text(
    aes(label = label, y = (ymin + ymax) / 2, vjust = if_else(is_thin, 0.7, 0.5)),
    color = "white", fontface = "bold", size = 3.4
  ) +
  geom_text(
    data = df_total,
    aes(x = factor(apc_year), y = grand_total, label = label),
    inherit.aes = FALSE,
    vjust = -0.6, fontface = "bold", size = 4, color = "black"
  ) +
  scale_fill_manual(values = pal_oa_types, name = NULL) +
  scale_y_continuous(
    labels = label_number(scale_cut = cut_short_scale(), prefix = "$"),
    expand = expansion(mult = c(0, 0.12))
  ) +
  labs(
    title = "Estimate of annual APC spend (in USD) by OA type",
    subtitle = "Adjusted for inflation to 2025 USD using CPI Advanced Economies.",
    caption = "Some publishers are not covered in every year (De Gruyter and OUP begin in 2021, IOP is missing 2020, and Sage's 2019 and 2025 data are gold-only and hybrid-only respectively).",
    x = NULL, y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor    = element_blank(),
    panel.grid.major.y = element_blank(),
    axis.text.x  = element_text(size = 12, face = "bold", color = "black"),
    axis.text.y  = element_blank(),
    axis.line.x.bottom  = element_line(color = "black", linewidth = 0.4),
    legend.position = "top",
    legend.justification = "left",
    plot.title    = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(color = "black", size = 11),
    plot.caption = element_textbox_simple(
      size = 9,
      color = "black",
      lineheight = 1.3,
      margin = margin(t = 12)
    )
  )

ggsave("figures/fig2_apc_spend_chicklet.png", width = 7, height = 5, dpi = 300, bg = "white")
```

### Numbers for plot description

Gold share of total spend across all years

```{r gold-share, echo=FALSE}
gold_share <- df_year_oa_type %>%
  group_by(oa_status) %>%
  summarise(total = sum(total), .groups = "drop") %>%
  mutate(total_m = total / 1e6,
         share = total / sum(total) * 100) %>%
  select(-total)

gold_share %>%
  mutate(total_m = sprintf("%.2f", total_m),
         share = sprintf("%.2f", share))
```

Growth 2019 → 2025, per OA type

```{r oa-type-growth, echo=FALSE}
growth <- df_year_oa_type %>%
  filter(apc_year %in% c(2019, 2025)) %>%
  select(apc_year, oa_status, total) %>%
  mutate(total_m = total / 1e6) %>%
  select(-total) %>%
  pivot_wider(names_from = apc_year, values_from = total_m, names_prefix = "y") %>%
  mutate(growth_pct = round((y2025 - y2019) / y2019 * 100, 2))

growth %>%
  mutate(across(c(y2019, y2025, growth_pct), ~ sprintf("%.2f", .x)))
```

## Publishers

```{r function-clean-publisher-labels,echo=FALSE}
# Clean publisher labels
clean_label <- function(x) {
  x %>%
    str_replace_all("_", " ") %>%
    str_to_title() %>%
    str_replace("Acs", "ACS") %>%
    str_replace("Cambridge University Press - Cup", "CUP") %>%
    str_replace("Edp", "EDP") %>%
    str_replace("Ieee", "IEEE") %>%
    str_replace("Iop", "IOP") %>%
    str_replace("Mdpi", "MDPI") %>%
    str_replace("Oup", "OUP") %>%
    str_replace("Plos", "PLOS")
}
```

```{r fig3-plot-function,echo=FALSE}
make_fig3 <- function(data, ends_data, title, y_breaks, pal, use_repel = TRUE) {
  lab_layer <- if (use_repel) {
    geom_text_repel(
      data = ends_data,
      aes(x = x_lab, y = revenue_m, label = lab),
      hjust              = 0,
      direction          = "y",
      size               = 3.1,
      fontface           = "bold",
      segment.colour     = "grey75",
      segment.size       = 0.25,
      min.segment.length = 0.4,
      box.padding        = 0.25,
      show.legend        = FALSE
    )
  } else {
    geom_text(
      data = ends_data,
      aes(x = x_lab, y = revenue_m, label = lab),
      hjust       = 0,
      size        = 3.1,
      fontface    = "bold",
      show.legend = FALSE
    )
  }

  ggplot(data, aes(apc_year, revenue_m, colour = publisher)) +

    # hand-drawn gridlines: stop at 2025, don't run into the label gutter
    geom_segment(
      data = tibble(y = y_breaks),
      aes(x = 2019, xend = 2025, y = y, yend = y),
      inherit.aes = FALSE,
      colour = "grey92", linewidth = 0.4
    ) +

    geom_line(linewidth = 0.9) +
    geom_point(size = 1.6) +
    lab_layer +

    scale_x_continuous(
      breaks = 2019:2025,
      expand = expansion(mult = c(0.03, 0.42))   # gutter for the labels
    ) +
    scale_y_continuous(
      breaks = y_breaks,
      labels = label_dollar(suffix = "M", big.mark = ",", accuracy = 1),
      expand = expansion(mult = c(0.05, 0.08))
    ) +
    scale_colour_manual(values = pal, drop = TRUE) +
    labs(title = title, x = NULL, y = NULL) +
    theme_minimal(base_size = 12) +
    theme(
      legend.position = "none",
      panel.grid      = element_blank(),
      plot.title      = element_text(face = "bold", size = 12,
                                     margin = margin(b = 6))
    )
}
```

### Figure 3. Estimate of annual APC spend (in USD) by publisher adjusted for inflation to 2025 USD using CPI Advanced Economies.

```{r fig3-publishers-apc-spend,echo=FALSE,fig.width = 6.5, fig.height = 7.5, fig.dpi=300}
library(scales)
library(patchwork)
library(ggrepel)

# Data 
rev_by_pub <- apc_aggregate %>%
  filter(oa_status %in% c("Gold OA", "Hybrid OA")) %>%
  mutate(revenue_adj = sum * to_2025[as.character(apc_year)]) %>%
  group_by(publisher, apc_year) %>%
  summarise(revenue_m = sum(revenue_adj, na.rm = TRUE) / 1e6, .groups = "drop") %>%
  mutate(publisher = clean_label(publisher)) %>%
  filter(!(publisher == "IOP" & apc_year == 2020)) # no data for 2020

# Rank publishers by total revenue across all years
top5 <- rev_by_pub %>%
  group_by(publisher) %>%
  summarise(tot = sum(revenue_m), .groups = "drop") %>%
  slice_max(tot, n = 5) %>%
  pull(publisher)

rev_by_pub <- rev_by_pub %>%
  mutate(group = if_else(publisher %in% top5, "Top 5 publishers", "Other publishers"))

# End-of-line labels, pinned just right of the last data point
ends <- rev_by_pub %>%
  group_by(publisher, group) %>%
  filter(apc_year == max(apc_year)) %>%
  ungroup() %>%
  mutate(
    lab = paste0(publisher, "  $",
                 formatC(round(revenue_m), format = "d", big.mark = ","),
                 "M"),
    x_lab = 2025.25
  )

# Plot 
p_top <- make_fig3(
  filter(rev_by_pub, group == "Top 5 publishers"),
  filter(ends,       group == "Top 5 publishers"),
  "Top 5 publishers",
  y_breaks  = seq(0, 1000, by = 250),
  pal       = pal_top5,
  use_repel = FALSE
)

p_rest <- make_fig3(
  filter(rev_by_pub, group == "Other publishers"),
  filter(ends,       group == "Other publishers"),
  "Next 9 publishers",
  y_breaks  = seq(0, 120, by = 30),
  pal = pal_other,
  use_repel = TRUE
)

publishers_apc_spend_total <- p_top / p_rest +
  plot_annotation(
    title    = "Estimate of annual APC spend (in USD) by publisher",
    subtitle = "Adjusted for inflation to 2025 USD using CPI Advanced Economies.",
    caption = "Sage's 2019 and 2025 figures reflect gold-only and hybrid-only data respectively.",
    theme    = theme(
      plot.title    = element_text(face = "bold", size = 15),
      plot.subtitle = element_text(colour = "black", size = 11,
                                   margin = margin(b = 10)),
      plot.caption = element_textbox_simple(
      size = 9,
      color = "black",
      lineheight = 1.3,
      margin = margin(t = 12)
    )
    )
  )

ggsave("figures/fig3_publishers_apc_spend_total.svg", publishers_apc_spend_total,
       width = 6.5, height = 7.5, dpi = 300, bg = "white")

publishers_apc_spend_total
```

(For the published figure, label positioning was adjusted with Adobe Illustrator)


#### Numbers for plot description

MDPI gold output, 2022–2024

```{r mdpi-output-change, echo=FALSE}
mdpi_output <- journals_pub_counts_apc %>%
  filter(publisher == "MDPI", year %in% 2022:2024, oa_status == "Gold OA") %>%
  group_by(year) %>%
  summarise(total_apcable = sum(goldDiamondHybrid, na.rm = TRUE), .groups = "drop") %>%
  mutate(chg_from_2022 = (total_apcable - total_apcable[year == 2022]) /
                          total_apcable[year == 2022] * 100)
mdpi_output
```

Frontiers spend trend, 2022–2024

```{r frontiers-trend, echo=FALSE}
frontiers_trend <- df_spend_by_publisher_oa_type_inflation %>%
  filter(publisher == "FRONTIERS", apc_year %in% 2022:2024) %>%
  group_by(apc_year) %>%
  summarise(total_m = sum(sum_adj, na.rm = TRUE) / 1e6, .groups = "drop") %>%
  arrange(apc_year) %>%
  mutate(pct_change_from_2022 = round((total_m - total_m[apc_year == 2022]) /
                                        total_m[apc_year == 2022] * 100, 1),
         yoy_pct_change = round((total_m - lag(total_m)) / lag(total_m) * 100, 1))

frontiers_trend %>%
  mutate(across(c(total_m, pct_change_from_2022, yoy_pct_change), ~ sprintf("%.2f", .x)))
```

MDPI, Frontiers, and Wiley gold output change, 2022–2023

```{r output-change-mdpi-frontiers-wiley, echo=FALSE}
output_change <- journals_pub_counts_apc %>%
  filter(publisher %in% c("MDPI", "FRONTIERS", "WILEY"),
         year %in% c(2022, 2023),
         oa_status == "Gold OA") %>%
  mutate(apcable = case_when(
    publisher %in% c("MDPI", "FRONTIERS") ~ goldDiamondHybrid,
    publisher == "WILEY" ~ gold,
    TRUE ~ NA_real_
  )) %>%
  group_by(publisher, year) %>%
  summarise(total_apcable = sum(apcable, na.rm = TRUE), .groups = "drop") %>%
  group_by(publisher) %>%
  summarise(
    y2022 = total_apcable[year == 2022],
    y2023 = total_apcable[year == 2023],
    pct_change = (y2023 - y2022) / y2022 * 100,
    .groups = "drop"
  )

output_change
```

IEEE spend growth, 2019–2025

```{r ieee-growth, echo=FALSE}
ieee_growth <- df_spend_by_publisher_oa_type_inflation %>%
  filter(publisher == "IEEE", apc_year %in% c(2019, 2025)) %>%
  group_by(apc_year) %>%
  summarise(total_m = sum(sum_adj, na.rm = TRUE) / 1e6, .groups = "drop") %>%
  pivot_wider(names_from = apc_year, values_from = total_m, names_prefix = "y") %>%
  mutate(growth_pct = round((y2025 - y2019) / y2019 * 100, 1))
ieee_growth
```

### Figure 4. Estimate of annual APC revenue (in USD) by publisher and OA type adjusted for inflation to 2025 USD using CPI Advanced Economies.

```{r fig4-publishers-apc-spend-oa-type, echo=FALSE, include=FALSE}
library(ggtext)
library(scales)
library(ggh4x)
library(grid)

# Data
df_spend_fig4 <- df_spend_by_publisher_oa_type_inflation %>%
    mutate(sum_adj = if_else(sum_adj == 0, NA_real_, sum_adj)) # so that Sage 2025 doesn't print as $0

# Explicit panel order (by apc spend in 2025)
pub_order <- c(
  "ELSEVIER", "SPRINGER_NATURE", "MDPI",
  "WILEY", "FRONTIERS", "OUP",
  "ACS", "PLOS", "CAMBRIDGE_UNIVERSITY_PRESS_-_CUP",
  "IOP", "SAGE", "IEEE",
  "DE_GRUYTER", "EDP"
)

# Fail on any name mismatch rather than silently dropping a panel
stopifnot(length(setdiff(pub_order, unique(df_spend_fig4$publisher))) == 0)
stopifnot(length(setdiff(unique(df_spend_fig4$publisher), pub_order)) == 0)

pub_labels <- setNames(clean_label(pub_order), pub_order)

df_spend_fig4_ordered <- df_spend_fig4 %>%
  mutate(publisher = factor(publisher, levels = pub_order))

# Peaks in panel order to check these clear the row ceilings set below
df_spend_fig4_ordered %>%
  group_by(publisher) %>%
  summarise(peak = max(sum_adj, na.rm = TRUE), .groups = "drop") %>%
  arrange(publisher) %>%
  print(n = Inf)

# Endpoint labels
df_ends <- df_spend_fig4_ordered %>%
  filter(apc_year == 2025) %>%
  mutate(label = label_number(scale_cut = cut_short_scale(),
                              prefix = "$", accuracy = 1)(sum_adj))

## Sage has no 2025 gold data; use its 2024 gold endpoint instead
df_ends_sage <- df_spend_fig4_ordered %>%
  filter(apc_year == 2024, publisher == "SAGE", oa_status == "Gold OA") %>%
  mutate(label = label_number(scale_cut = cut_short_scale(),
                              prefix = "$", accuracy = 1)(sum_adj))

df_ends <- bind_rows(df_ends, df_ends_sage) %>%
  mutate(publisher = factor(publisher, levels = pub_order))

# Per-row y-axis ceilings 
nc <- 3   

# Row 1: Elsevier, Springer Nature, MDPI
# Row 2: Wiley, Frontiers, OUP
# Row 3: ACS, PLOS, CUP
# Row 4: IOP, SAGE, IEEE
# Row 5: De Gruyter, EDP
row_specs <- list(
  list(limit = 700e6, breaks = c(0, 200e6, 400e6, 600e6)),
  list(limit = 320e6, breaks = c(0, 100e6, 200e6, 300e6)),
  list(limit =  60e6, breaks = c(0,  20e6,  40e6,  60e6)),
  list(limit =  60e6, breaks = c(0,  20e6,  40e6,  60e6)),
  list(limit =  8e6, breaks = c(0,   4e6,   8e6,  12e6))
)

y_scales <- vector("list", length(pub_order))

for (row_num in seq_along(row_specs)) {
  positions <- ((row_num - 1) * nc + 1):(row_num * nc)
  positions <- positions[positions <= length(pub_order)]
  spec <- row_specs[[row_num]]
  for (pos in positions) {
    y_scales[[pos]] <- scale_y_continuous(
      breaks = spec$breaks,
      labels = label_number(scale_cut = cut_short_scale(), prefix = "$"),
      limits = c(0, spec$limit),
      expand = expansion(mult = c(0.05, 0.15))
    )
  }
}

# OUP sits in row 2 (with Wiley and Frontiers) but belongs on a 60M axis
oup_pos <- which(pub_order == "OUP")
y_scales[[oup_pos]] <- scale_y_continuous(
  breaks = c(0, 20e6, 40e6, 60e6),
  labels = label_number(scale_cut = cut_short_scale(), prefix = "$"),
  limits = c(0, 60e6),
  expand = expansion(mult = c(0.05, 0.15))
)

# Plot
p <- ggplot(
  df_spend_fig4_ordered,
  aes(
    x = apc_year,
    y = sum_adj,
    colour = oa_status,
    group = oa_status
  )
) +
  geom_line(linewidth = 0.85) +
  geom_point(size = 3, shape = 16) +
  geom_text(
    data = df_ends,
    aes(label = label),
    hjust = -0.15,
    size = 5,
    fontface = "bold",
    show.legend = FALSE
  ) +
  facet_wrap2(
    ~ publisher,
    scales = "free_y",
    ncol = nc,
    labeller = as_labeller(pub_labels)
  ) +
  facetted_pos_scales(y = y_scales) +
  scale_colour_manual(values = pal_oa_types, name = NULL) +
  scale_x_continuous(
    breaks = c(2019, 2022, 2025),
    expand = expansion(mult = c(0.05, 0.42))   # room for the endpoint labels
  ) +
  labs(
    title    = "Estimate of annual APC spend (in USD) by publisher and OA type",
    subtitle = "Adjusted for inflation to 2025 USD using CPI Advanced Economies.",
    caption  = "Panels are ordered by total 2025 APC spend Note that the y-axis scale differs between panels.",
    x = NULL,
    y = NULL
  ) +
  theme_minimal() +
  theme(
    panel.grid.major.x   = element_blank(),
    panel.grid.major.y   = element_line(colour = "grey92", linewidth = 0.3),
    panel.grid.minor     = element_blank(),
    strip.text           = element_text(
      face = "bold",
      size = 14,
      colour = "black"
    ),
    legend.position      = "top",
    legend.justification = "left",
    legend.text          = element_text(size = 16),
    legend.key.width     = unit(2, "lines"),
    axis.text.x          = element_text(size = 14),
    axis.text.y          = element_text(size = 14),
    plot.title           = element_text(face = "bold", size = 18),
    plot.subtitle        = element_text(colour = "black", size = 16),
    plot.caption         = element_text(
      size = 14,
      colour = "black",
      margin = margin(t = 15),
      hjust = 0
    ),
    plot.caption.position = "plot",
    plot.margin = margin(b = 15)
  )
```

```{r fig4-publishers-apc-spend-oa-type-plot, echo=FALSE, fig.width=10, fig.height=13}
print(p)

png(
  "figures/fig4_apc_spend_publisher_oa_type.png",
  width = 10 * 300,
  height = 13 * 300,
  res = 300
)
print(p)
grid.lines(
  x  = unit(0.67, "npc"),
  y  = unit(c(0.55, 0.72), "npc"),
  gp = gpar(
    col = "black",
    lwd = 2.5,
    lty = "dashed"
  )
)
dev.off()

```

### Table 2. Growth of article output and APC spend (adjusted for inflation to 2025 USD using CPI Advanced Economies) from 2019 to 2025 per publisher and OA type.

```{r table2-spend-growth, echo=FALSE}
library(kableExtra)

# Spend growth
spend_growth <- df_spend_by_publisher_oa_type_inflation %>%
  filter(apc_year %in% c(2019, 2025))

## growth rate per publisher, split by OA status
growth_by_type <- spend_growth %>%
  group_by(publisher, oa_status, apc_year) %>%
  summarise(total = sum(sum_adj), .groups = "drop") %>%
  pivot_wider(names_from = apc_year, values_from = total, names_prefix = "y") %>%
  mutate(growth_pct = round((y2025 - y2019) / y2019 * 100, 1))

## growth rate per publisher, Gold+Hybrid combined
growth_combined <- spend_growth %>%
  group_by(publisher, apc_year) %>%
  summarise(total = sum(sum_adj), .groups = "drop") %>%
  pivot_wider(names_from = apc_year, values_from = total, names_prefix = "y") %>%
  mutate(oa_status = "Gold+Hybrid",
         growth_pct = round((y2025 - y2019) / y2019 * 100, 1))

## the "all publishers" row, split by OA status
growth_all <- spend_growth %>%
  group_by(oa_status, apc_year) %>%
  summarise(total = sum(sum_adj), .groups = "drop") %>%
  pivot_wider(names_from = apc_year, values_from = total, names_prefix = "y") %>%
  mutate(publisher = "all publishers",
         growth_pct = round((y2025 - y2019) / y2019 * 100, 1))

## the "all publishers" row, Gold+Hybrid combined
growth_all_combined <- spend_growth %>%
  group_by(apc_year) %>%
  summarise(total = sum(sum_adj), .groups = "drop") %>%
  pivot_wider(names_from = apc_year, values_from = total, names_prefix = "y") %>%
  mutate(publisher = "all publishers", oa_status = "Gold+Hybrid",
         growth_pct = round((y2025 - y2019) / y2019 * 100, 1))

# spend growth table
table2_spend_growth <- bind_rows(growth_by_type, growth_combined,
                                  growth_all, growth_all_combined) %>%
  select(publisher, oa_status, growth_pct) %>%
  pivot_wider(names_from = oa_status, values_from = growth_pct) %>%
  select(publisher, spend_gh = `Gold+Hybrid`, spend_g = `Gold OA`, spend_h = `Hybrid OA`)

# Article output growth
## article-count growth per publisher, split by OA status
article_growth <- journals_pub_counts_apc %>%
  filter(year %in% c(2019, 2025), oa_status %in% c("Gold OA", "Hybrid OA")) %>%
  group_by(publisher, oa_status, year) %>%
  summarise(total = sum(goldDiamondHybrid, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = year, values_from = total, names_prefix = "y") %>%
  mutate(growth_pct = round((y2025 - y2019) / y2019 * 100, 1))

## article-count growth per publisher, Gold+Hybrid combined
article_growth_combined <- journals_pub_counts_apc %>%
  filter(year %in% c(2019, 2025), oa_status %in% c("Gold OA", "Hybrid OA")) %>%
  group_by(publisher, year) %>%
  summarise(total = sum(goldDiamondHybrid, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = year, values_from = total, names_prefix = "y") %>%
  mutate(oa_status = "Gold+Hybrid",
         growth_pct = round((y2025 - y2019) / y2019 * 100, 1))

## the "all publishers" row, split by OA status
article_growth_all <- journals_pub_counts_apc %>%
  filter(year %in% c(2019, 2025), oa_status %in% c("Gold OA", "Hybrid OA")) %>%
  group_by(oa_status, year) %>%
  summarise(total = sum(goldDiamondHybrid, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = year, values_from = total, names_prefix = "y") %>%
  mutate(publisher = "all publishers",
         growth_pct = round((y2025 - y2019) / y2019 * 100, 1))

## the "all publishers" row, Gold+Hybrid combined
article_growth_all_combined <- journals_pub_counts_apc %>%
  filter(year %in% c(2019, 2025), oa_status %in% c("Gold OA", "Hybrid OA")) %>%
  group_by(year) %>%
  summarise(total = sum(goldDiamondHybrid, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = year, values_from = total, names_prefix = "y") %>%
  mutate(publisher = "all publishers", oa_status = "Gold+Hybrid",
         growth_pct = round((y2025 - y2019) / y2019 * 100, 1))

# article grwoth table
table2_article_growth <- bind_rows(article_growth, article_growth_combined,
                                    article_growth_all, article_growth_all_combined) %>%
  select(publisher, oa_status, growth_pct) %>%
  pivot_wider(names_from = oa_status, values_from = growth_pct) %>%
  select(publisher, articles_gh = `Gold+Hybrid`, articles_g = `Gold OA`, articles_h = `Hybrid OA`)

# Combine article + spend growth into one table
table2_full <- table2_article_growth %>%
  left_join(table2_spend_growth, by = "publisher") %>%
  mutate(publisher = if_else(publisher == "all publishers",
                              "All publishers",
                              clean_label(publisher))) %>%
  arrange(publisher != "All publishers", publisher)

# Add n/a for Sage due to data gaps
table2_full <- table2_full %>%
  mutate(across(-publisher, ~ if_else(str_detect(publisher, regex("sage", ignore_case = TRUE)), NA_real_, .x)))

# Save
write_csv(table2_full, "table2_growth_full.csv")

# Print
table2_full %>%
  mutate(across(-publisher, ~ if_else(is.na(.x), "n/a", sprintf("%+.1f%%", .x)))) %>%
  knitr::kable(
    col.names = c("Publisher", "Gold+Hybrid", "Gold", "Hybrid",
                  "Gold+Hybrid", "Gold", "Hybrid"),
    align = "lrrrrrr",
    caption = "Table 2. Growth of article output and APC spend (adjusted for inflation to 2025 USD using CPI Advanced Economies) from 2019 to 2025 per publisher and OA type."
  ) %>%
  kable_styling(full_width = FALSE) %>%
  add_header_above(c(
    " " = 1,
    "Number of APC-able publications" = 3,
    "Spend estimate based on APCs in USD (adjusted to 2025 prices)" = 3
  ))
```

## APC listed vs. paid

```{r data-fig5-6-7, echo=FALSE, include=FALSE}
# Derive the revenue ranking from the spend data
rev_2025 <- apc_aggregate %>%
  filter(oa_status %in% c("Gold OA", "Hybrid OA"), apc_year == 2025) %>%
  group_by(publisher) %>%
  summarize(revenue = sum(sum, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(revenue)) %>%
  mutate(rev_rank = row_number())

print(rev_2025, n = Inf)   # confirm the top five

N_TOP <- 5

# Load the APC data
df_2025 <- journals_pub_counts_apc %>%
  filter(apc_year == 2025, !is.na(apc_usd), apc_usd >= 0)

# LIST = the listed APCs One observation per journal.
list_data <- df_2025 %>%
  select(publisher, oa_status, apc_usd) %>%
  mutate(type = "List")

# PAID = the realized distribution. One observation per article. $0 journals retained
paid_data <- df_2025 %>%
  filter(goldDiamondHybrid > 0) %>%
  uncount(goldDiamondHybrid) %>%
  select(publisher, oa_status, apc_usd) %>%
  mutate(type = "Paid")

# Function clean labels
clean_pub <- function(x) {
  x <- if_else(x == "CAMBRIDGE_UNIVERSITY_PRESS_-_CUP", "CUP", x)
  recode(x,
    "SPRINGER_NATURE" = "Springer Nature", "DE_GRUYTER" = "De Gruyter",
    "ELSEVIER" = "Elsevier", "WILEY" = "Wiley", "FRONTIERS" = "Frontiers", "SAGE" = "Sage")
}

# Join the ranking on the raw publisher strings, then clean
plot_df <- bind_rows(list_data, paid_data) %>%
  filter(oa_status != "unknown") %>%
  left_join(rev_2025 %>% select(publisher, rev_rank), by = "publisher") %>%
  mutate(
    publisher = clean_pub(publisher),
    group      = if_else(!is.na(rev_rank) & rev_rank <= N_TOP, "Top5", "Next9"),
    type      = fct_relevel(type, "List", "Paid"),
    facet_lab = paste0(publisher, " \u00b7 ", oa_status)
  )
```

### Figure 5. Distribution of the number of journals per listed APC (journal level) and the number of articles per paid APC (article level) for 2025 by OA type, all publishers. Medians are indicated with a dashed line.

```{r fig5-apc-listed-paid-overall, echo=FALSE, include=FALSE}
library(colorspace)
library(scales)
library(tidysdm)
library(ggtext)

# Paid = one row per article, $0 journals retained on both sides.

# Pool across all publishers, one panel per OA status 
pooled_df <- plot_df %>%
  mutate(oa_status = fct_relevel(oa_status, "Gold OA", "Hybrid OA"))

# Medians, with dashed leader lines out to the labels
pooled_meds <- pooled_df %>%
  group_by(oa_status, type) %>%
  summarize(median_val = median(apc_usd), .groups = "drop") %>%
  mutate(
    x_start   = 1,
    x_end     = if_else(type == "List", 0.55, 1.45),
    x_lab     = if_else(type == "List", 0.52, 1.48),
    x_pt      = 1 + if_else(type == "List", -0.01, 0.01),
    hjust_val = if_else(type == "List", 1, 0),
    label_txt = label_dollar(accuracy = 1)(median_val),
    pt_fill   = if_else(type == "List",
                        lighten("#541F3F", 0.45),
                        lighten("#A85A47", 0.45))
  )
```

```{r fig5-apc-listed-paid-overall-plot, echo=FALSE, fig.dpi=300}
ggplot(pooled_df, aes(x = "Profile", y = apc_usd, fill = type)) +
  geom_split_violin(width = 0.7, alpha = 0.9, colour = "grey25",
                    linewidth = 0.3, adjust = 1.4) +
  # dashed leader from the spine out to the label
  geom_segment(
    data = pooled_meds,
    aes(x = x_start, xend = x_end, y = median_val, yend = median_val),
    inherit.aes = FALSE,
    linetype = "dashed", colour = "grey15", linewidth = 0.5
  ) +
  geom_point(
    data = pooled_meds,
    aes(x = x_pt, y = median_val, fill = I(pt_fill)),
    inherit.aes = FALSE,
    shape = 21, size = 2.8, colour = "grey20", stroke = 0.5,
    show.legend = TRUE
  ) +
  geom_text(
    data = pooled_meds,
    aes(x = x_lab, y = median_val, label = label_txt, hjust = hjust_val),
    inherit.aes = FALSE,
    size = 4, fontface = "bold", colour = "grey15"
  ) +
  facet_wrap(vars(oa_status), nrow = 1) +
  scale_y_continuous(
    labels = label_comma(),
    breaks = c(seq(0, 12000, by = 2000), 13000),
    expand = expansion(mult = c(0.03, 0.05))
  ) +
  coord_cartesian(ylim = c(0, 13000)) +
  scale_x_discrete(expand = expansion(mult = 0.85)) +
  scale_fill_manual(values = c("List" = "#541F3F", "Paid" = "#A85A47")) +
  labs(
    title    = "Listed and paid APCs (in USD), 2025: all publishers pooled",
    subtitle = "The median article costs more than the median journal in both gold and hybrid OA.",
    x = NULL, y = NULL, fill = NULL,
    caption  = "Left half: listed prices, one per journal. Right half: weighted by article volume. Dashed lines mark medians."
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.text = element_text(size = 11),
    legend.position       = "top",
    legend.justification  = "left",
    plot.title            = element_textbox_simple(
                              face = "bold", size = 14,
                              margin = margin(b = 6), lineheight = 1.2),
    plot.subtitle         = element_textbox_simple(
                              size = 10.5, colour = "black",
                              margin = margin(t = 2, b = 4), lineheight = 1.3),
    plot.caption          = element_textbox_simple(
                              size = 10.5, colour = "black",
                              margin = margin(t = 10), lineheight = 1.3),
    plot.title.position   = "plot",
    plot.caption.position = "plot",
    strip.text            = element_text(face = "bold", size = 11),
    strip.background      = element_blank(),
    panel.grid.minor      = element_blank(),
    panel.grid.major.x    = element_blank(),
    panel.spacing         = unit(1.5, "lines"),
    axis.text.x           = element_blank(),
    axis.ticks.x          = element_blank()
  )

ggsave("figures/fig5-pooled.svg", width = 8, height = 5.5, dpi = 600, bg = "white")
```

(For the published figure, median dots in the legend were edited using Adobe Illustrator to match color in the figure)

### Figure 6 and 7. Distribution of listed (journal-level, blue) and paid (article-level, orange) APCs by publishers and OA type, 2025.

```{r function-fig6-7, echo=FALSE}
make_fig6_7 <- function(group_code, levels_vec, title, subtitle, caption,
                           ncol_facets, y_max = 7000) {

  keep <- shift_df %>% filter(group == group_code)
  if (nrow(keep) == 0) {
    warning("No publishers in group '", group_code, "'.")
    return(NULL)
  }

  d <- plot_df %>%
    filter(facet_lab %in% levels_vec) %>%
    mutate(facet_lab = factor(facet_lab, levels = levels_vec))

  meds <- keep %>%
    select(facet_lab, list_med, paid_med) %>%
    pivot_longer(c(list_med, paid_med), names_to = "type", values_to = "median_val") %>%
    mutate(
      type      = if_else(type == "list_med", "List", "Paid"),
      type      = fct_relevel(type, "List", "Paid"),
      pt_fill   = if_else(type == "List",
                        lighten("#541F3F", 0.45),
                        lighten("#A85A47", 0.45)),
      facet_lab = factor(facet_lab, levels = levels_vec)
    )

  ggplot(d, aes(x = "Profile", y = apc_usd, fill = type)) +
    geom_split_violin(width = 0.5, alpha = 0.85, colour = "grey25",
                      linewidth = 0.3, adjust = 1.6) +
    geom_point(
      data = meds,
      aes(x = "Profile", y = median_val, fill = I(pt_fill)),
      inherit.aes = FALSE,
      shape = 21, size = 2, colour = "grey20", stroke = .7
    ) +
    facet_wrap(vars(facet_lab), ncol = ncol_facets) +
    scale_y_continuous(
      labels = label_dollar(scale = 1e-3, suffix = "k", accuracy = 1),
      limits = c(0, y_max),
      breaks = seq(0, y_max, by = 3000),
      expand = expansion(mult = c(0.05, 0.08))
    ) +
    scale_x_discrete(expand = expansion(mult = 0.6)) +
    scale_fill_manual(values = c("List" = "#541F3F", "Paid" = "#A85A47")) +
    labs(title = title, subtitle = subtitle, caption = caption,
         x = NULL, y = NULL, fill = NULL) +
    theme_minimal(base_size = 12) +
    theme(
      plot.margin = margin(t = 5),
      legend.position       = "top",
      legend.justification  = "left",
      plot.title            = element_textbox_simple(
                                face = "bold", size = 14,
                                margin = margin(b = 6), lineheight = 1.2),
      plot.subtitle         = element_textbox_simple(
                                size = 10.5, colour = "black",
                                margin = margin(t = 2, b = 12), lineheight = 1.3),
      plot.caption = element_textbox_simple(
        size = 8, colour = "black",
        margin = margin(t = 10), lineheight = 1.3
        ),
      plot.title.position   = "plot",
      plot.caption.position = "plot",
      strip.text = ggtext::element_textbox_simple(
  face = "bold", size = 9, halign = 0.5,
  lineheight = 1.3,
  margin = margin(t = 6, b = 6)
),
      strip.background      = element_blank(),
      panel.grid.minor      = element_blank(),
      panel.grid.major.x    = element_blank(),
      panel.spacing         = unit(1.0, "lines"),
      axis.text.x           = element_blank(),
      axis.ticks.x          = element_blank()
    )
}
```

```{r fig6-7-apc-listed-paid-by-publisher, include=FALSE}
library(scales)
library(tidysdm)
library(ggtext)

# Fig 6 & 7: Listed vs paid APCs, split by publisher revenue group

# Shift size and direction , per publisher x OA status cell
shift_df <- plot_df %>%
  group_by(publisher, oa_status, facet_lab, group, rev_rank) %>%
  summarize(
    list_med = median(apc_usd[type == "List"]),
    paid_med = median(apc_usd[type == "Paid"]),
    n_prices = n_distinct(apc_usd[type == "List"]),
    share_at = mean(apc_usd[type == "Paid"] == median(apc_usd[type == "List"])),
    .groups  = "drop"
  ) %>%
  filter(
    !is.na(paid_med),
    # SAGE Gold 2025 was not collected; the journals that do appear were picked up incidentally via other publishers' APC lists.
    !(publisher == "Sage" & oa_status == "Gold OA")
  ) %>%
  mutate(
    shift_usd = paid_med - list_med,
    shift_direction = case_when(
      shift_usd > 0 ~ "up",
      shift_usd < 0 ~ "down",
      TRUE          ~ "mirror"
    ),
    shift_direction = fct_relevel(shift_direction, "up", "mirror", "down")
  )

# check claim that all Top5 publishers shift up and more varied pattern for smaller publishers
shift_df %>% count(group, shift_direction) %>% print()

cat("\nTop5, all cells:\n")
shift_df %>%
  filter(group == "Top5") %>%
  arrange(rev_rank, oa_status) %>%
  select(publisher, oa_status, list_med, paid_med, shift_usd, shift_direction) %>%
  print(n = Inf)

cat("\nNext9, by shift direction:\n")
shift_df %>%
  filter(group == "small") %>%
  arrange(shift_direction, desc(shift_usd)) %>%
  select(publisher, oa_status, list_med, paid_med, shift_usd, shift_direction) %>%
  print(n = Inf)

# Panel ordering
# Top5: by shift size.
# Next9: by shift direction (up, mirror, down), then by shift size within shift direction
lev_top5 <- shift_df %>%
  filter(group == "Top5") %>%
  arrange(desc(shift_usd)) %>%
  pull(facet_lab)

lev_next9 <- shift_df %>%
  filter(group == "Next9") %>%
  arrange(shift_direction, desc(shift_usd), publisher) %>%
  pull(facet_lab)

# y-axis max
max_by_panel <- plot_df %>%
  filter(type == "List") %>%
  group_by(publisher, oa_status, facet_lab, group) %>%
  summarize(
    n_journals = n(),
    min_apc    = min(apc_usd),
    med_apc    = median(apc_usd),
    max_apc    = max(apc_usd),
    .groups    = "drop"
  ) %>%
  # attach the panel order used in each figure
  mutate(
    panel_pos = case_when(
      group == "Top5" ~ match(facet_lab, lev_top5),
      group == "Next9" ~ match(facet_lab, lev_next9)
    ),
    figure = if_else(group == "Top5", "4a (Top5)", "4b (Next9)"),
    clipped = max_apc > 7000
  ) %>%
  filter(!is.na(panel_pos)) %>%      # drop cells excluded by the n_prices filter
  arrange(group, panel_pos) %>%
  select(figure, panel_pos, facet_lab, n_journals,
         min_apc, med_apc, max_apc, clipped)

print(max_by_panel, n = Inf)

# Plot
p_top5 <- make_fig6_7(
  "Top5", lev_top5,
  "Listed and paid APCs (in USD), 2025: Top 5 publishers by APC spend",
  "In every case the median article costs more than the median journal: most articles appear in the more expensive part of the portfolio.",
  "Panels ordered by size of the gap. Left half: listed prices, one per journal. Right half: weighted by article volume. Points mark medians. Journals above $9,000 not shown: 20 at Elsevier (1.1% of hybrid titles, 0.8% of hybrid article volume) and 39 at Springer Nature (1.9% of hybrid titles, 4.8% of hybrid volume).",
  ncol_facets = 4,
  y_max = 9000
)

p_next9 <- make_fig6_7(
  "Next9", lev_next9,
  "Listed and paid APCs (in USD), 2025: Next 9 publishers",
  "Article volume shifts up the price list at some publishers, tracks it at others, and shifts down at the rest.",
  "Panels grouped by direction of shift, then ordered by size of the gap. Left half: listed prices, one per journal. Right half: weighted by article volume. Points mark medians. Four journals above $6,000 not shown (3 hybrid at OUP, 1 gold at PLOS; under 1% of article volume at each).",
  ncol_facets = 4,
  y_max = 6000
)

ggsave("figures/fig6-top5.svg", p_top5, width = 6.5, height = 6,   dpi = 300, bg = "white")
ggsave("figures/fig7-next9.svg", p_next9, width = 6.5, height = 8.5, dpi = 300, bg = "white")

```

```{r fig6-print, echo=FALSE, fig.dpi=300,fig.height=6, fig.width=6.5}
p_top5
```

```{r fig7-print, echo=FALSE, fig.dpi=300,fig.height=8.5, fig.width=6.5}
p_next9

```

(For the published figures, median dots in the legend were edited using Adobe Illustrator to match color in the figure)

### Table 3. Median of 2025 APCs listed (journal level) and paid (article level) per OA status per publisher.

```{r table3-apc-listed-paid, echo=FALSE}
library(kableExtra)
library(scales)

# Assumes plot_df is already built (List = journal level, Paid = article level,
# $0 journals retained on both sides).

# Per-publisher medians
tbl_pub <- plot_df %>%
  group_by(publisher, oa_status, type) %>%
  summarize(med = median(apc_usd), .groups = "drop")

# All-publishers row: pooled across publishers
tbl_all <- plot_df %>%
  group_by(oa_status, type) %>%
  summarize(med = median(apc_usd), .groups = "drop") %>%
  mutate(publisher = "all publishers")

# Assemble wide
table3 <- bind_rows(tbl_all, tbl_pub) %>%
  mutate(
    oa    = if_else(str_detect(oa_status, "Gold"), "gold", "hybrid"),
    col   = paste0(oa, "_", tolower(type))
  ) %>%
  select(publisher, col, med) %>%
  pivot_wider(names_from = col, values_from = med) %>%
  select(publisher,
         gold_list, gold_paid,
         hybrid_list, hybrid_paid) %>%
  # display formatting: dollars, "n/a" where the publisher has no such cell
  mutate(across(-publisher,
                ~ if_else(is.na(.x), "n/a", label_dollar(accuracy = 1)(.x)))) %>%
  arrange(publisher != "all publishers", publisher)   # all-publishers row first

# Save
write_csv(table3, "table3_listed_vs_paid_medians.csv")

# Print
table3 %>%
  knitr::kable(
    col.names = c("Publisher", "Listed", "Paid", "Listed", "Paid"),
    align = "lrrrr",
    caption = "Table 3. Median of 2025 APCs listed (journal level) and paid (article level) per OA status per publisher."
  ) %>%
  kable_styling(full_width = FALSE) %>%
  add_header_above(c(
    " " = 1,
    "Median gold APC 2025 (in USD)" = 2,
    "Median hybrid APC 2025 (in USD)" = 2
  ))

```
