---
title: "roara_apc_paper_publication_counts"
author: "Lisa Matthias"
date: "2026-07-29"
output: github_document
always_allow_html: true
---
This supplement provides the number of APC-able publications underlying the spend estimates reported in the main text.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

```{r include=FALSE}
library(tidyverse)
```

```{r data, include=FALSE}
article_counts <- read_csv("./data/article_counts_by_publisher_year_OAtype.csv")
```

```{r colors, include=FALSE}
# Palette OA Types
pal_oa_types <- c("Gold OA" = "#D6A94E", "Hybrid OA" = "#B85428")

# Palette Publishers
pal_top5 <- c(
  "Elsevier"        = "#ff7f0f",  
  "Springer Nature" = "#00316e",   
  "MDPI"            = "#697754",   
  "Wiley"           = "#b680bf",   
  "Frontiers"       = "#069494"    
)

pal_other <- c(
  "OUP"        = "#205f4b", 
  "ACS"        = "#7B7BE8",
  "PLOS"       = "#7e1a2f",
  "CUP"        = "#ee8577",
  "IOP"        = "#4A8FBF",
  "Sage"       = "#B3B177",
  "IEEE"       = "#2E4057",
  "De Gruyter" = "#a06177",
  "EDP"        = "#D58E40"
)
```

### Table S3.1 Number of APC-able publications per OA type, 2019–2025.

```{r tableS3-1-article-count-overall,echo=FALSE}
library(kableExtra)

# Get article counts
pub_count_table <- article_counts %>%
  group_by(apc_year) %>%
  summarise(Gold = sum(`Gold OA`, na.rm = TRUE),
            Hybrid = sum(`Hybrid OA`, na.rm = TRUE),
            .groups = "drop")

# Add 2019-2025 total row
pub_count_totals <- pub_count_table %>%
  summarise(across(c(Gold, Hybrid), sum, na.rm = TRUE)) %>%
  mutate(apc_year = "2019-2025")

pub_count_table_full <- pub_count_table %>%
  mutate(apc_year = as.character(apc_year)) %>%
  bind_rows(pub_count_totals)

# Print
pub_count_table_full %>%
  mutate(Gold = scales::comma(Gold),
         Hybrid = scales::comma(Hybrid)) %>%
  knitr::kable(
    col.names = c("Publication year", "Gold", "Hybrid"),
    align = "lrr",
    caption = "Table S3.1 Number of APC-able publications per OA type, 2019–2025."
  ) %>%
  kable_styling(full_width = FALSE)
```

## OA type
### Figure S3.2 Number of APC-able publications per OA type, 2019–2025.

```{r figS3-2-articles-gold-hybrid-chicklet, echo=FALSE, fig.dpi=300}
library(ggchicklet)   
library(scales)
library(ggtext)

df_articles_by_publisher_oa_type <- article_counts %>%
  pivot_longer(cols = c(`Gold OA`, `Hybrid OA`),
               names_to = "oa_status", values_to = "n_articles")

# Total articles per year, per OA type (summed across all publishers)
df_year_oa_type_articles <- df_articles_by_publisher_oa_type %>%
  group_by(apc_year, oa_status) %>%
  summarise(total = sum(n_articles, na.rm = TRUE), .groups = "drop")

# Lock stacking order explicitly: Gold OA bottom, Hybrid OA top
df_year_oa_type_articles <- df_year_oa_type_articles %>%
  mutate(oa_status = factor(oa_status, levels = c("Gold OA", "Hybrid OA")))

# Compute cumulative stack positions manually so labels can be placed precisely
df_year_oa_type_articles <- df_year_oa_type_articles %>%
  arrange(apc_year, oa_status) %>%
  group_by(apc_year) %>%
  mutate(
    ymax        = cumsum(total),
    ymin        = ymax - total,
    grand_total_bar = sum(total),
    seg_share   = total / grand_total_bar
  ) %>%
  ungroup()

df_year_oa_type_articles <- df_year_oa_type_articles %>%
  mutate(is_thin = seg_share < 0.12)

# Label format: mixed K/M style, matching the main paper's format_mixed()
format_articles_mixed <- function(x) {
  if_else(
    x >= 1e6,
    paste0(round(x / 1e6, 1), "M"),
    paste0(scales::comma(round(x / 1e3, 0)), "K")
  )
}

# Endpoint labels: total per segment, placed inside/at top of each chicklet
df_year_oa_type_articles <- df_year_oa_type_articles %>%
  mutate(label = format_articles_mixed(total))

# Grand total per year (for the overall annotation above each bar)
df_total_articles <- df_year_oa_type_articles %>%
  group_by(apc_year) %>%
  summarise(grand_total = sum(total, na.rm = TRUE)) %>%
  mutate(label = format_articles_mixed(grand_total))

# Plot
ggplot(df_year_oa_type_articles, aes(x = factor(apc_year), y = total, fill = oa_status)) +
  geom_chicklet(
    radius = grid::unit(3, "pt"),
    width = 0.7
  ) +
  geom_text(
    aes(label = label, y = (ymin + ymax) / 2, vjust = if_else(is_thin, 0.7, 0.5)),
    color = "white", fontface = "bold", size = 3.4
  ) +
  geom_text(
    data = df_total_articles,
    aes(x = factor(apc_year), y = grand_total, label = label),
    inherit.aes = FALSE,
    vjust = -0.6, fontface = "bold", size = 4, color = "black"
  ) +
  scale_fill_manual(values = pal_oa_types, name = NULL) +
  scale_y_continuous(
    labels = label_number(scale_cut = cut_short_scale()),
    expand = expansion(mult = c(0, 0.12))
  ) +
  labs(
    title = "Number of APC-able publications by OA type",
    caption = "Some publishers are not covered in every year (De Gruyter and OUP begin in 2021, IOP is missing 2020, and Sage's 2019 and 2025 data are gold-only and hybrid-only respectively).",
    x = NULL, y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor    = element_blank(),
    panel.grid.major.y = element_blank(),
    axis.text.x  = element_text(size = 12, face = "bold", color = "black"),
    axis.text.y  = element_blank(),
    axis.line.x.bottom  = element_line(color = "black", linewidth = 0.4),
    legend.position = "top",
    legend.justification = "left",
    plot.title    = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(color = "black", size = 11),
    plot.caption = element_textbox_simple(
      size = 9,
      color = "black",
      lineheight = 1.3,
      margin = margin(t = 12)
    )
  )

ggsave("figures/figS3_2_articles_chicklet.png", width = 7, height = 5, dpi = 300, bg = "white")
```

## Publisher
### Figure S3.3 Number of APC-able publications by publisher, 2019–2025.

```{r function-clean-publisher-labels,echo=FALSE}
# Clean publisher labels
clean_label <- function(x) {
  x %>%
    str_replace_all("_", " ") %>%
    str_to_title() %>%
    str_replace("Acs", "ACS") %>%
    str_replace("Cambridge University Press - Cup", "CUP") %>%
    str_replace("Edp", "EDP") %>%
    str_replace("Ieee", "IEEE") %>%
    str_replace("Iop", "IOP") %>%
    str_replace("Mdpi", "MDPI") %>%
    str_replace("Oup", "OUP") %>%
    str_replace("Plos", "PLOS")
}
```

```{r function-make-fig-count,echo=FALSE}
library(tidyverse)
library(scales)
library(ggrepel)

make_fig_count <- function(data, ends_data, title, y_breaks, pal, use_repel = TRUE,
                            label_fn = scales::comma) {

  lab_layer <- if (use_repel) {
    geom_text_repel(
      data = ends_data,
      aes(x = x_lab, y = value, label = lab),
      hjust              = 0,
      direction          = "y",
      size               = 3.1,
      fontface           = "bold",
      segment.colour     = "grey75",
      segment.size       = 0.25,
      min.segment.length = 0.4,
      box.padding        = 0.25,
      show.legend        = FALSE
    )
  } else {
    geom_text(
      data = ends_data,
      aes(x = x_lab, y = value, label = lab),
      hjust       = 0,
      size        = 3.1,
      fontface    = "bold",
      show.legend = FALSE
    )
  }

  ggplot(data, aes(apc_year, value, colour = publisher)) +

    # hand-drawn gridlines: stop at 2025, don't run into the label gutter
    geom_segment(
      data = tibble(y = y_breaks),
      aes(x = 2019, xend = 2025, y = y, yend = y),
      inherit.aes = FALSE,
      colour = "grey92", linewidth = 0.4
    ) +

    geom_line(linewidth = 0.9) +
    geom_point(size = 1.6) +
    lab_layer +

    scale_x_continuous(
      breaks = 2019:2025,
      expand = expansion(mult = c(0.03, 0.42))   # gutter for the labels
    ) +
    scale_y_continuous(
      breaks = y_breaks,
      labels = label_fn,
      expand = expansion(mult = c(0.05, 0.08))
    ) +
    scale_colour_manual(values = pal, drop = TRUE) +
    labs(title = title, x = NULL, y = NULL) +
    theme_minimal(base_size = 12) +
    theme(
      legend.position = "none",
      panel.grid      = element_blank(),
      plot.title      = element_text(face = "bold", size = 12,
                                     margin = margin(b = 6))
    )
}
```

```{r figS3-3-publishers-article-output,echo=FALSE,fig.width = 6.5, fig.height = 7.5, fig.dpi=300}
library(patchwork)

articles_by_pub <- article_counts %>%
  pivot_longer(cols = c(`Gold OA`, `Hybrid OA`),
               names_to = "oa_status", values_to = "n_articles") %>%
  group_by(publisher, apc_year) %>%
  summarise(value = sum(n_articles, na.rm = TRUE), .groups = "drop") %>%
  mutate(publisher = clean_label(publisher)) %>%
  filter(!(publisher == "IOP" & apc_year == 2020))

top5_articles <- articles_by_pub %>%
  group_by(publisher) %>%
  summarise(tot = sum(value), .groups = "drop") %>%
  slice_max(tot, n = 5) %>%
  pull(publisher)

articles_by_pub <- articles_by_pub %>%
  mutate(group = if_else(publisher %in% top5_articles, "Top 5 publishers", "Other publishers"))

ends_articles <- articles_by_pub %>%
  group_by(publisher, group) %>%
  filter(apc_year == max(apc_year)) %>%
  ungroup() %>%
  mutate(
    lab = paste0(publisher, "  ", formatC(round(value), format = "d", big.mark = ",")),
    x_lab = 2025.25
  )

p_top5_articles <- make_fig_count(
  filter(articles_by_pub, group == "Top 5 publishers"),
  filter(ends_articles,   group == "Top 5 publishers"),
  "Top 5 publishers",
  y_breaks = seq(0, 350000, by = 70000),   # covers max ~344,287
  pal      = pal_top5,
  use_repel = FALSE
)

p_next9_articles <- make_fig_count(
  filter(articles_by_pub, group == "Other publishers"),
  filter(ends_articles,   group == "Other publishers"),
  "Next 9 publishers",
  y_breaks  = seq(0, 35000, by = 7000),   # covers max ~34,156
  pal       = pal_other,
  use_repel = TRUE
)

publishers_article_output <- p_top5_articles / p_next9_articles +
  plot_annotation(
    title    = "Number of APC-able publications by publisher",
    caption  = "Sage's 2019 and 2025 figures reflect gold-only and hybrid-only data respectively.",
    theme    = theme(
      plot.title    = element_text(face = "bold", size = 15),
      plot.subtitle = element_text(colour = "black", size = 11,
                                   margin = margin(b = 10)),
      plot.caption  = element_textbox_simple(
        size = 9,
        color = "black",
        lineheight = 1.3,
        margin = margin(t = 12)
      )
    )
  )

ggsave("figures/figS3_3_publishers_article_output.svg", publishers_article_output,
       width = 6.5, height = 7.5, dpi = 300, bg = "white")

publishers_article_output
```

### Figure S3.4 Number of APC-able publications by publisher and OA type, 2019–2025.

```{r figS3-4-publishers-article-count-oa-type, echo=FALSE, include=FALSE}
library(ggtext)
library(scales)
library(ggh4x)
library(grid)

# Data
df_articles_fig <- df_articles_by_publisher_oa_type %>%
  group_by(publisher, apc_year, oa_status) %>%
  summarise(n_articles = sum(n_articles, na.rm = TRUE), .groups = "drop") %>%
  # Explicitly remove SAGE 2025 Gold OA to prevent it from plotting
  filter(!(publisher == "SAGE" & apc_year == 2025 & oa_status == "Gold OA")) %>%
  mutate(n_articles = if_else(n_articles == 0, NA_real_, n_articles))

# Explicit panel order (matching Figure 4's publisher order)
pub_order <- c(
  "ELSEVIER", "SPRINGER_NATURE", "MDPI",
  "WILEY", "FRONTIERS", "OUP",
  "PLOS", "IEEE", "ACS", 
  "IOP", "CAMBRIDGE_UNIVERSITY_PRESS_-_CUP", "SAGE",
  "DE_GRUYTER", "EDP"
)

stopifnot(length(setdiff(pub_order, unique(df_articles_fig$publisher))) == 0)
stopifnot(length(setdiff(unique(df_articles_fig$publisher), pub_order)) == 0)

pub_labels <- setNames(clean_label(pub_order), pub_order)

df_articles_fig_ordered <- df_articles_fig %>%
  mutate(publisher = factor(publisher, levels = pub_order))

# Peaks in panel order — check these against the row ceilings set below
df_articles_fig_ordered %>%
  group_by(publisher) %>%
  summarise(peak = max(n_articles, na.rm = TRUE), .groups = "drop") %>%
  arrange(publisher) %>%
  print(n = Inf)

# Endpoint labels
df_ends <- df_articles_fig_ordered %>%
  filter(apc_year == 2025) %>%
  mutate(label = label_number(scale_cut = cut_short_scale(), accuracy = 1)(n_articles))

## Sage has no 2025 gold data; use its 2024 gold endpoint instead
df_ends_sage <- df_articles_fig_ordered %>%
  filter(apc_year == 2024, publisher == "SAGE", oa_status == "Gold OA") %>%
  mutate(label = label_number(scale_cut = cut_short_scale(), accuracy = 1)(n_articles))

df_ends <- bind_rows(df_ends, df_ends_sage) %>%
  mutate(publisher = factor(publisher, levels = pub_order))

# Per-row y-axis ceilings 
nc <- 3

# Run this first to set real ceilings:
df_articles_fig_ordered %>% group_by(publisher) %>% summarise(peak = max(n_articles, na.rm=TRUE))

row_specs <- list(
  list(limit = 300000, breaks = c(0, 100000, 200000, 300000)),           # Row 1: Elsevier, SN, MDPI
  list(limit = 100000, breaks = c(0, 25000, 50000, 75000, 100000)),      # Row 2: Wiley, Frontiers, OUP
  list(limit =  25000, breaks = c(0,  5000, 10000, 15000, 20000, 25000)),# Row 3: ACS, PLOS, CUP
  list(limit =  16000, breaks = c(0,  4000,  8000, 12000, 16000)),       # Row 4: IOP, SAGE, IEEE
  list(limit =   8000, breaks = c(0,  2000,  4000,  6000,  8000))        # Row 5: De Gruyter, EDP
)

y_scales <- vector("list", length(pub_order))

for (row_num in seq_along(row_specs)) {
  positions <- ((row_num - 1) * nc + 1):(row_num * nc)
  positions <- positions[positions <= length(pub_order)]
  spec <- row_specs[[row_num]]
  for (pos in positions) {
    y_scales[[pos]] <- scale_y_continuous(
      breaks = spec$breaks,
      labels = label_number(scale_cut = cut_short_scale()),
      limits = c(0, spec$limit),
      expand = expansion(mult = c(0.05, 0.15))
    )
  }
}

# OUP override — much smaller than Wiley/Frontiers in its row (peak 21,499)
oup_pos <- which(pub_order == "OUP")
y_scales[[oup_pos]] <- scale_y_continuous(
  breaks = c(0, 5000, 10000, 15000, 20000, 25000),
  labels = label_number(scale_cut = cut_short_scale()),
  limits = c(0, 25000),
  expand = expansion(mult = c(0.05, 0.15))
)

# IEEE override — largest in its row (peak 18,995), 16,000 ceiling would clip it
ieee_pos <- which(pub_order == "IEEE")
y_scales[[ieee_pos]] <- scale_y_continuous(
  breaks = c(0, 5000, 10000, 15000, 20000, 25000),
  labels = label_number(scale_cut = cut_short_scale()),
  limits = c(0, 25000),
  expand = expansion(mult = c(0.05, 0.15))
)

# ACS override — largest in its row (peak 18,995), 16,000 ceiling would clip it
acs_pos <- which(pub_order == "ACS")
y_scales[[acs_pos]] <- scale_y_continuous(
  breaks = c(0, 4000, 8000, 12000, 16000),
  labels = label_number(scale_cut = cut_short_scale()),
  limits = c(0, 16000),
  expand = expansion(mult = c(0.05, 0.15))
)

# Plot
p_articles <- ggplot(
  df_articles_fig_ordered,
  aes(
    x = apc_year,
    y = n_articles,
    colour = oa_status,
    group = oa_status
  )
) +
  geom_line(linewidth = 0.85) +
  geom_point(size = 3, shape = 16) +
  geom_text(
    data = df_ends,
    aes(label = label),
    hjust = -0.15,
    size = 5,
    fontface = "bold",
    show.legend = FALSE
  ) +
  facet_wrap2(
    ~ publisher,
    scales = "free_y",
    ncol = nc,
    labeller = as_labeller(pub_labels)
  ) +
  facetted_pos_scales(y = y_scales) +
  scale_colour_manual(values = pal_oa_types, name = NULL) +
  scale_x_continuous(
    breaks = c(2019, 2022, 2025),
    expand = expansion(mult = c(0.05, 0.42))
  ) +
  labs(
    title    = "Number of APC-able publications by publisher and OA type",
    caption  = "Panels are ordered by total number of APC-able publications. Note that the y-axis scale differs between panels.",
    x = NULL,
    y = NULL
  ) +
  theme_minimal() +
  theme(
    panel.grid.major.x   = element_blank(),
    panel.grid.major.y   = element_line(colour = "grey92", linewidth = 0.3),
    panel.grid.minor     = element_blank(),
    strip.text           = element_text(face = "bold", size = 14, colour = "black"),
    legend.position      = "top",
    legend.justification = "left",
    legend.text          = element_text(size = 16),
    legend.key.width     = unit(2, "lines"),
    axis.text.x          = element_text(size = 14),
    axis.text.y          = element_text(size = 14),
    plot.title           = element_text(face = "bold", size = 18),
    plot.subtitle        = element_text(colour = "black", size = 16),
    plot.caption         = element_text(
      size = 14, colour = "black", margin = margin(t = 15), hjust = 0
    ),
    plot.caption.position = "plot",
    plot.margin = margin(b = 15)
  )
```

```{r figS3-4-publishers-article-count-oa-type-plot, echo=FALSE, fig.width=10, fig.height=13}
print(p_articles)

png(
  "figures/figS3_4_article_count_publisher_oa_type.png",
  width = 10 * 300,
  height = 13 * 300,
  res = 300
)
print(p_articles)
dev.off()
```

(For the published version, dashed lines were added with Adobe Illustrator)